### 2014-03-28 vignette/msnid_vignette.Rnw
mostly addition of vignette

%\VignetteIndexEntry{MSnID Package for Handling MS/MS Identifications}
% \VignetteDepends{BiocStyle, msmsTests}
%\VignetteKeywords{Documentation}
%\VignettePackage{MSnID}
\documentclass[11pt]{article}

<<style, eval=TRUE, echo=FALSE, results=tex>>=
BiocStyle::latex()
@

\usepackage[authoryear,round]{natbib}

\title{\Rpackage{MSnID} Package for Handling MS/MS Identifications}
\author{Vladislav A. Petyuk}

\begin{document}
\SweaveOpts{concordance=TRUE, eval=TRUE, prefix.string=graphics}

\maketitle
\tableofcontents

\section{Introduction}
The \Rpackage{MSnID} package is a convenience tool to manipulating
MS/MS identifications in R/Bioconductor environment. The input file 
can be a flat text file with the results of MS/MS search in a table
format or mzIdentML files. The main S4 class for storing data is
\Rclass{MSnID}. A class for handling MS/MS filtering criteria
is \Rclass{MSnIDFilter}.

\section{Case study}
<<init, results=hide>>=
library("MSnID")
@

\subsection{Downloading example mzIdentML files}
As an example we will use datasets from C.elegans proteomics study describing the effect of daf-16 mutation on the protein abundances \cite{Depuydt2013,Depuydt2014}. There will be 5 datasets corresponding to long-living glp-4; daf-2 strain and 5 datasets corresponding to the strain with compensatory mutation in daf-16 transcription factor that restores (shortens) life-span back to normal. Files available at PeptideAtlas repository \href{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00308}{PASS00308}.
<<>>=
try(setInternet2(FALSE),silent=TRUE)
ftp <- "ftp://PASS00308:PJ5348t@ftp.peptideatlas.org/"
meta <- read.delim(sprintf("%sSample_Metadata.txt", ftp), as.is=T)
meta <- subset(meta, Age == 'young' & Diet == 'ff') # explore the effect of daf-16.
#for( dataset in meta$PNNL.Dataset.Name){
#   cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
#                       ftp, dataset)
#   download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
#}
@

\subsection{Setting up working directory}
First, the \Robject{MSnID} object has to be initialized. The main argument is
path to the working directory. This directory will be used for storing cached 
analysis results. Caching/memoisation mechanism is based on \CRANpkg{R.cache}.
<<>>=
library("MSnID")
# start the project by providing work directory
msnid <- MSnID(".")
@

\subsection{Reading MS/MS data}
The main intended way to read MS/MS results is by parsing mzIdentML files (*.mzid or *.mzid.gz extensions) using \Biocpkg{mzID} package facilities.
<<>>=
mzids <- list.files(".", pattern=".mzid.gz")
# in this case I read all *.mzid.gz files present in the current directory.
msnid <- read_mzIDs(msnid, mzids)
@
Alternative way is to pass MS/MS search results as peptide-to-spectra matches (PSMs) \Rclass{data.frame}. Internally PSMs stored as \CRANpkg{data.table} object.
<<eval=FALSE>>=
psms(msnid) <- yourPSMresults
@

\subsection{Updating columns}
Note, to have a fully functioning \Robject{MSnID} object, the following columns has to be present.
<<>>=
MSnID:::.mustBeColumns
@
Check what are the current column names of the MS/MS search results.
<<>>=
names(msnid)
@
Trivial updates to some of the columns to make sure upper/lower letter cases are correct.
<<>>=
msnid$Accession <- msnid$accession
msnid$isDecoy <- msnid$isdecoy
msnid$calculatedMassToCharge <- msnid$calculatedmasstocharge
msnid$experimentalMassToCharge <- msnid$experimentalmasstocharge
msnid$chargeState <- msnid$chargestate
msnid$spectrumID <- msnid$spectrumid
@
Creating "Peptide" column as peptide sequence with flanking amino acids (X.XXXX.X).
<<>>=
msnid$Peptide <- paste(msnid$pre, msnid$pepseq, msnid$post, sep='.')
@
Removing the columns we don't need anymore.
<<>>=
msnid$accession <- NULL
msnid$pepseq <- NULL
msnid$pre <- NULL
msnid$post <- NULL
msnid$isdecoy <- NULL
msnid$spectrumid <- NULL
msnid$experimentalmasstocharge <- NULL
msnid$calculatedmasstocharge <- NULL
msnid$chargestate <- NULL
@

\subsection{Exploring the MSnID object}
Printing the \Robject{MSnID} object returns some basic information such as
\begin{itemize}
   \item Working directory.
   \item Number of spectrum files used to generate data.
   \item Number of peptide-to-spectrum matches and corresponding FDR.
   \item Number of unique peptide sequences and corresponding FDR.
   \item Number of unique proteins or amino acid sequence accessions and corresponding FDR.
\end{itemize}
False discovery rate or FDR is defined here as a ratio of hits to decoy accessions to the non-decoy accessions. In terms of forward and revese protein sequences that would mean ratio of \#reverse/\#forward. While computing FDRs of PSMs and unique peptide sequences is trivial, definition of protein (accession) FDR is a subject for discussion in the field of proteomics. Here, protein (accession) FDR is computed the same way as in IDPicker software \cite{Zhang2007} and simply constitutes a ratio of unqiue accessions from decoy component to non-decoy component of sequence database.
<<>>=
show(msnid)
@

\subsection{Analysis of peptide sequences}
A particular properties of peptide sequences we are interested in are 1) irregular cleavages at the termini of the peptides and 2) missing cleavage site within the peptide sequences. The default regular expressions of valid and missed cleavage patterns correspond to trypsin.
<<>>=
# creates new column NumIrregCleavages
msnid <- assess_termini(msnid, validCleavagePattern="[RK]\\.[^P]")
# creates new column NumMissCleavages
msnid <- assess_missed_cleavages(msnid, missedCleavagePattern="[KR](?=[^P$])")
@
Now the object has two more columns, NumIrregCleavages and NumMissCleavages, evidently corresponding to the number of termini with irregular cleavages and number of missed cleavages within the peptide sequence.
<<label=missedCleavages, fig=TRUE, include=FALSE, width=9>>=
pepCleav <- unique(psms(msnid)[,c("NumMissCleavages", "isDecoy", "Peptide")])
pepCleav <- as.data.frame(table(pepCleav[,c("NumMissCleavages", "isDecoy")]))
library("ggplot2")
ggplot(pepCleav, aes(x=NumMissCleavages, y=Freq, fill=isDecoy)) + 
   geom_bar(stat='identity', position='dodge') + 
   ggtitle("Number of Missed Cleavages")
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-missedCleavages}
\end{center}
\pagebreak[0]
Peptide sequences can by accessed by directly using \Rcode{\$} operator. For example:
<<label=lengths, fig=TRUE, include=FALSE, width=9>>=
# counting number of cysteins per peptide sequence
msnid$NumCys <- sapply(lapply(strsplit(msnid$Peptide,''),'==','C'),sum)
# calculating peptide lengths
msnid$PepLength <- nchar(msnid$Peptide) - 4
pepLen <- unique(psms(msnid)[,c("PepLength", "isDecoy", "Peptide")])
ggplot(pepLen, aes(x=PepLength, fill=isDecoy)) + 
   geom_histogram(position='dodge', binwidth=3) +
   ggtitle("Distribution on of Peptide Lengths")
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-lengths}
\end{center}
\pagebreak[0]

\subsection{Trimming the data}
The main facility for trimming or filtering the data is \Rfunction{apply\_filter} 
function. The second argument can be either 1) a string representing expression that will be evaluated in the context of data.frame containing MS/MS results or 2) \Rclass{MSnFilter} class object (explained below). Note, the reduction in FDR. True identifications tend to be fully tryptic and contain fewer missed cleavages.
class object.
<<>>=
# take a look at original FDRs
show(msnid)
# Leave only fully tryptic.
msnid <- apply_filter(msnid, "NumIrregCleavages == 0")
show(msnid)
# Retain peptides with at most 2 missed cleavages.
msnid <- apply_filter(msnid, "NumMissCleavages <= 2")
show(msnid)
@

\subsection{Parent ion mass measurement accuracy}
Assuming both \Rcode{calculatedMassToCharge} and \Rcode{experimentalMassToCharge} are present in \Rcode{names(msnid)}, one can access parent ion mass measurement in points per million (ppm) units.
<<label=ppmOriginal, fig=TRUE, include=FALSE, width=9>>=
ppm <- mass_measurement_error(msnid)
ggplot(as.data.frame(ppm), aes(x=ppm)) + geom_histogram(binwidth=100)
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-ppmOriginal}
\end{center}
\pagebreak[0]
Note, although the MS/MS search was done with +/- 20ppm parent ion mass tolerance, error stretch over 1000 in ppm units. The reason is that setting of the MS/MS search engine MS-GF+ assumed imperfect peak picking for fragmentation and considered peptides that were +/- 1 Da (\textsuperscript{13}C-\textsuperscript{12}C to be exact) off.
<<label=deltaMass, fig=TRUE, include=FALSE, width=9>>=
dM <- with(psms(msnid), (experimentalMassToCharge-calculatedMassToCharge)*chargeState)
x <- data.frame(dM, isDecoy=msnid$isDecoy)
ggplot(x, aes(x=dM, fill=isDecoy)) + geom_histogram(position='stack')
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-deltaMass}
\end{center}
\pagebreak[0]
Selection and reporting of wrong monoisotopic masses can be corrected.
<<label=ppmCorrectedMass, fig=TRUE, include=FALSE, width=9>>=
msnid.fixed <- correct_peak_selection(msnid)
ppm <- mass_measurement_error(msnid.fixed)
ggplot(as.data.frame(ppm), aes(x=ppm)) + geom_histogram(binwidth=0.25)
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-ppmCorrectedMass}
\end{center}
\pagebreak[0]
Alternatively, one can simply apply a filter to remove any matches that do not fit the +/- 20 ppm tolerance.
<<label=ppmFiltered20, fig=TRUE, include=FALSE, width=9>>=
msnid.chopped <- apply_filter(msnid, "abs(mass_measurement_error(msnid)) < 20")
ppm <- mass_measurement_error(msnid.chopped)
ggplot(as.data.frame(ppm), aes(x=ppm)) + geom_histogram(binwidth=0.25)
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-ppmFiltered20}
\end{center}
\pagebreak[0]
!NOTE! Make sure the text is correct.
For further processing we'll consider \Rcode{msnid.chopped} data that is adjusted for 1 Da errors in picking of wrong monoisotopic peak. Note, if the center of the histogram is significantly shifted from zero, \Rcode{experimentalMassToCharge} can be recalibrated.
<<label=ppmRecalibrated, fig=TRUE, include=FALSE, width=9>>=
msnid <- recalibrate(msnid.chopped)
ppm <- mass_measurement_error(msnid)
ggplot(as.data.frame(ppm), aes(x=ppm)) + geom_histogram(binwidth=0.25)
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-ppmRecalibrated}
\end{center}
\pagebreak[0]

\subsection{Defining and optimizing filters for MS/MS identifications}
First, let's define criteria we are going to use for data filtration. First will be based on MS-GF+ Spectrum E-value, second will be absolute mass measurement error of the parent ion.
<<label=msmsScoreDistribution, fig=TRUE, include=FALSE, width=9>>=
msnid$msmsScore <- -log10(msnid$`ms-gf:specevalue`)
params <- psms(msnid)[,c("msmsScore","isDecoy")]
ggplot(params) + geom_density(aes(x = msmsScore, color = isDecoy, ..count..))
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-msmsScoreDistribution}
\end{center}
\pagebreak[0]
<<label=absPpmDistribution, fig=TRUE, include=FALSE, width=9>>=
msnid$absParentMassErrorPPM <- abs(mass_measurement_error(msnid))
params <- psms(msnid)[,c("absParentMassErrorPPM","isDecoy")]
ggplot(params) + geom_density(aes(x = absParentMassErrorPPM, color = isDecoy, ..count..))
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-absPpmDistribution}
\end{center}
\pagebreak[0]
MS/MS fiters are handled by a special \Rclass{MSnIDFilter} class objects. Individual filtering criteria can be set by name (that is present in \Rcode{names(msnid)}), comparison operator (>, <, = , ...) defining what to retain, and a threshold value.
<<>>=
filtObj <- MSnIDFilter(msnid)
filtObj$absParentMassErrorPPM <- list(comparison="<", threshold=10.0)
filtObj$msmsScore <- list(comparison=">", threshold=7.0)
# print filter
show(filtObj)
# evaluate filter
evaluate_filter(msnid, filtObj, level="PSM")
evaluate_filter(msnid, filtObj, level="Peptide")
evaluate_filter(msnid, filtObj, level="Accession")
@
The threshold values in the example above are not necessarily optimal and set just be in the range of probable values. Filters can be optimized to ensure maximum number of identifications (peptide-to-spectrum matches, unique peptide sequences or proteins) within given FDR upper limit.
<<>>=
# brute-force optimization by enumeration all the parameter combinations
# these should be good starting parameters for follow-up optimizations
system.time({
   filtObj.grid <- optimize_filter(filtObj, msnid, fdr.max=0.01, 
                                  method="Grid", level="Peptide", n.iter=500)})
show(filtObj.grid)
# (absParentMassErrorPPM < 2) & (msmsScore > 7.8)

# Nelder-Mead optimization
set.seed(0)
system.time({
   filtObj.nm <- optimize_filter(filtObj.grid, msnid, fdr.max=0.01, 
                          method="Nelder-Mead", level="Peptide", n.iter=500)})
show(filtObj.nm)
# (absParentMassErrorPPM < 3) & (msmsScore > 7.8)

# simulated annealing optimization
set.seed(0)
system.time({
   filtObj.sann <- optimize_filter(filtObj.grid, msnid, fdr.max=0.01, 
                                  method="SANN", level="Peptide", n.iter=500)})
show(filtObj.sann)
# (absParentMassErrorPPM < 2.2) & (msmsScore > 7.6)
@
Let's check which filter turned out to be the best. The differece between the optimized filters is likely to be very small.
<<>>=
evaluate_filter(msnid, filtObj, level="Peptide")
evaluate_filter(msnid, filtObj.grid, level="Peptide")
evaluate_filter(msnid, filtObj.nm, level="Peptide")
evaluate_filter(msnid, filtObj.sann, level="Peptide")
@
We'll choose and apply filter produced by simulated annealing optimization procedure. Identifications that matched to decoy and contaminant protein sequences can be removed by providing filters in the forms of text strings that will be evaluated in the context of PSM table.
<<>>=
# apply the optimized filter
msnid <- apply_filter(msnid, filtObj.sann)
show(msnid)
# let's remove reverse/decoy and Contaminants
msnid <- apply_filter(msnid, "!isDecoy")
show(msnid)
msnid <- apply_filter(msnid, "!grepl('Contaminant',Accession)")
show(msnid)
@

\subsection{Downstream quantitative analysis}
The \Rpackage{MSnID} package is aimed at providing convenience functionality to handle MS/MS identifications. Quantitation \textit{per se} is outside of the scope of the package. The only type of quantitation that can be seamlessly blended with MS/MS identification analysis is so-called \emph{spectral counting} approah. In such an approach a peptide abundance is considered to be directly proportional to the number of matched MS/MS spectra. In its turn protein abunance is proportional to the sum of the number of spectra of the matching peptides. The \Rclass{MSnID} object can be converted to \Rclass{MSnSet} object form \Biocpkg{MSnbase} that extends generic Bioconductor \Rclass{eSet} to quantitative proteomics data.
<<label=convertingToMSnSet>>=
msnset <- as(msnid, "MSnSet")
# Note, feature data is peptide-centric. Peptide to protein assigments
# stored in feature data.
head(fData(msnset))
# Note, sample names in msnset are based on file names that were used as input
# MS/MS search engine. Let's trim the file names to make them compatible with
# dataset names.
head(sampleNames(msnset))
head(meta)
# Update sample names. 
# Retain only dataset name portion from spectrum file names, 
# that were used as sample names
sampleNames(msnset) <- sub("(.*)_dta\\.txt","\\1",sampleNames(msnset))
show(msnset)
# prepare pheno data
rownames(meta) <- meta$PNNL.Dataset.Name
meta <- meta[,c("Letter.Replicate","Daf.16.type")]
pData(msnset) <- meta[sampleNames(msnset),]
validObject(msnset)
@
Currently the features in the \Rclass{MSnSet} object are peptides. However, we plan to quantify differences at level of relative protein abundances. Peptides can be rolled-up to protein level using \Rfunction{combineFeatures} function from \Biocpkg{MSnbase}. Ambiguity in peptide-to-protein assignment is handled by \Rcode{redundancy.handler} argument. It can allow ambiguity of peptide assignment to \emph{multiple} proteins or \emph{unique} mapping only by discarding peptide mapping two or more proteins.
<<label=assessingRedundancy, fig=TRUE, include=FALSE>>=
# assessing the extent of peptide/protein mapping redundancy problem
redundancy <- table(sapply(fData(msnset)$Accession, length), dnn="redundancy")
redundancy <- 100 * prop.table(redundancy)
ggplot(as.data.frame(redundancy), aes(x=factor(1), y=Freq, fill=redundancy)) +
   geom_bar(stat='identity', width=1) + 
   coord_polar(theta='y') + 
   xlab('') + ylab('') +
   labs(fill='Redundancy') +
   scale_x_discrete(breaks = NULL)
# original number of features/peptides in the data
length(featureNames(msnset))
@
\begin{center}
\includegraphics[width=0.8\textwidth]{graphics-assessingRedundancy}
\end{center}
\pagebreak[4]
<<label=rollingToProteinLevel>>==
# summing allowing peptide to contribute multiple proteins
msnset.prot.1 <- combineFeatures(msnset, fData(msnset)$Accession, 
                                  redundancy.handler="multiple", 
                                  fun="sum", cv=FALSE)
# summing using only non-redundant peptides.
# Note, in this case 1/3 of peptides will be discarded
msnset.prot.2 <- combineFeatures(msnset, fData(msnset)$Accession, 
                                  redundancy.handler="unique", 
                                  fun="sum", cv=FALSE)
# Subset to proteins that present in at least 6 samples out of 10.
msnset <- msnset.prot.2[rowSums(exprs(msnset.prot.2) > 0) >= 6,]
# check how many proteins left
length(featureNames(msnset))
@
For statistical testing we will leverage functionality available in \Biocpkg{msmsTests}. Here we are comparing the difference between \textit{glp-4; daf-2} long-living strain with a strain that bears compensatory mutation within \textit{daf-16} transcription factor. For statistical testing of the count data we will use quasi-Poisson model \Rfunction{msms.glm.qlll}.
<<label=statisticalTests, fig=TRUE, include=FALSE>>=
library("msmsTests")
alt.f <- "y ~ Daf.16.type + 1"
null.f <- "y ~ 1"
div <- colSums(exprs(msnset))
res <- msms.glm.qlll(msnset, alt.f, null.f, div=div)
sum(p.adjust(res$p.value, "BH") < 0.05)
# Post-test filter
lst <- test.results(res,msnset,pData(msnset)$Daf.16.type,"wt","mut",div,
                    alpha=0.05,minSpC=0,minLFC=1,
                    method="BH")
res.volcanoplot(lst$tres, min.LFC=1, max.pval=0.05, ylbls=NULL, maxy=4)
@
\begin{center}
\includegraphics[width=0.4\textwidth]{graphics-statisticalTests}
\end{center}
\pagebreak[0]

\bibliographystyle{plainnat}
\bibliography{msnid}
\end{document}

### 2014-03-28 R/AllClasses.R
updates of c_elegans.R example

@@ -1,5 +1,7 @@
 
 
+.PROTON_MASS <- 1.007276466812
+
 setClass(Class="MSnID",
          representation(
             workDir="character", # working directory

### 2014-03-28 R/MSnID-methods.R

@@ -323,7 +323,8 @@ MSnID <- function(workDir='.', cleanCache=FALSE)
 
 .mustBeColumns <- c("Peptide", "Accession", "isDecoy", 
                     "calculatedMassToCharge",
-                    "experimentalMassToCharge", 
+                    "experimentalMassToCharge",
+                    "chargeState",
                     "spectrumFile", "spectrumID")
 
 setGeneric("psms",

### 2014-03-28 examples/c_elegans.R

@@ -20,6 +20,10 @@ msnid <- read_mzIDs(msnid, mzids)
 #---------------------------------------
 
 
+
+
+
+
 #--- CHECKING WHAT IS INSIDE -----------
 head(psms(msnid))
 names(msnid)
@@ -40,6 +44,7 @@ msnid$Accession <- msnid$accession
 msnid$isDecoy <- msnid$isdecoy
 msnid$calculatedMassToCharge <- msnid$calculatedmasstocharge
 msnid$experimentalMassToCharge <- msnid$experimentalmasstocharge
+msnid$chargeState <- msnid$chargestate
 msnid$spectrumID <- msnid$spectrumid
 # tidying up. removing left-over columns
 msnid$accession <- NULL
@@ -50,21 +55,23 @@ msnid$idecoy <- NULL
 msnid$spectrumid <- NULL
 msnid$experimentalmasstocharge <- NULL
 msnid$calculatedmasstocharge <- NULL
+msnid$chargestate <- NULL
 #-------------------------------------
 show(msnid)
-
-
 
 
 # --- EXTRA INFO ABOUT PEPTIDE SEQUENCES ---
 msnid <- assess_termini(msnid, validCleavagePattern="[RK]\\.[^P]")
 msnid <- assess_missed_cleavages(msnid, missedCleavagePattern="[KR](?=[^P$])")
-# visualize distributions of missed cleavages and non-tryptic termini
-temp <- psms(msnid)[,c("NumMissCleavages", "NumIrregCleavages", "Peptide")]
-temp <- unique(temp)
-barplot(table(temp$NumMissCleavages), main="Number of Missed Cleavages")
-barplot(table(temp$NumIrregCleavages), main="Distribution of Non-Tryptic Termini")
-#------------------------------------------
+# visualize distributions of missed cleavages as an example
+library("ggplot2")
+pepCleav <- unique(psms(msnid)[,c("NumMissCleavages", "isDecoy", "Peptide")])
+pepCleav <- as.data.frame(table(pepCleav[,c("NumMissCleavages", "isDecoy")]))
+p <- ggplot(pepCleav, aes(x=NumMissCleavages, y=Freq, fill=isDecoy))
+p + geom_bar(stat='identity', position='dodge')
+#----------------------------------------
+
+
 
 
 # --- TRIM THE DATA -----------------------
@@ -78,7 +85,7 @@ show(msnid)
 #-----------------------------------------
 
[…]

### 2014-04-01 R/AllClasses.R
fifth commit

@@ -6,24 +6,11 @@ setClass(Class="MSnID",
          representation(
             workDir="character", # working directory
             psms="data.table" # peptide-to-spectra matches
-#             filter="MSnIDFilter" # keeps record of the applied filter
             )
          )
 
 
 
-
-# Why does it have to be a seprate slot? Too complex for one?
-# Or perhaps, filter manipulations imply that it exists as a separate class. Does it?
-# 1) retrieving as string
-# 2) setting/updating particular pars...
-# ..a) how would you call a particular parameter? By name? *************
-# e.g. msnidObj$MSnFilter$ppm.abs$threshold <- 10 ? Not sure about this way of accessing
-# h.a? set_filter inteface? It is better, because there are multiple paths depending 
-# on what is provided.
-# Filter can exist within MSnIDObject, but get filter should return string (as well).
-# Because I apply it as string.
-# Do I need it as number?
 
 # partitions and parameter transforms to be added to filter class
 setClass(Class="MSnIDFilter",
@@ -33,140 +20,3 @@ setClass(Class="MSnIDFilter",
          )
 )
 
-
-# #######  BELOW ARE THE THOUGHTS ################################################
-# 
-# 
-# # should be able to set it like
-# # filtObj <- set_filter(filtObj, name="abs.ppm", comparison="<", threshold=2)
-# # filtObj <- set_filter(filtObj, name="abs.ppm") # acceptable too
-# # filtObj <- set_filter(filtObj, name="MSGFScore") # acceptable too
-# # filtObj <- optimize_filter_threshold( msnidObj, filtObj, FDR=0.05, level="peptide")
-# # eval_filter (at given level)? to get FDR and number of peptides
-# # msnid <- apply_filter( "MSnID", "character" | "MSnIDFilter")
-# # filtStr <- as(filtObj, "character")
-# 
-# 
-# # alternative
-# # set_filter(msnidObj, name="abs.ppm")
-# # set_filter(msnidObj, name="XCorr", append=TRUE)
-# # set_filter(msnidObj, name="msmsScore", append=FALSE)
-# # how about if I want to update the threshold is it append/replace
-# # action = c("append", "reset", "update")?
-# # how about rule:
-# # if not exist - append,
-# # if exist - update threshold (leave comparison if exists)
-# # reset is separate function
-# 
[…]

### 2014-04-01 R/AllGenerics.R

@@ -5,21 +5,23 @@
 setGeneric("set_filter",
            function(.Object, ...) standardGeneric("set_filter"))
 
-setGeneric("delete_psm_parameter",
-           function(.Object, ...) standardGeneric("delete_psm_parameter"))
+# setGeneric("delete_psm_parameter",
+#            function(.Object, ...) standardGeneric("delete_psm_parameter"))
 
-setGeneric("get_psm_parameter",
-           function(.Object, parName) standardGeneric("get_psm_parameter"))
+# setGeneric("get_psm_parameter",
+#            function(.Object, parName) standardGeneric("get_psm_parameter"))
 
-setGeneric("set_psm_parameter",
-           function(.Object, ...) standardGeneric("set_psm_parameter"))
+# setGeneric("set_psm_parameter",
+#            function(.Object, ...) standardGeneric("set_psm_parameter"))
 
 setGeneric("assess_missed_cleavages",
-           function(.Object, ...) standardGeneric("assess_missed_cleavages"))
+           function(.Object, missedCleavagePattern="[KR](?=[^P$])") 
+              standardGeneric("assess_missed_cleavages"))
 
 # this may be an odd way to handle assess_termini
 setGeneric("assess_termini", 
-           function(.Object, ...) standardGeneric("assess_termini"))
+           function(.Object, validCleavagePattern="[KR]\\.[^P]") 
+              standardGeneric("assess_termini"))
 
 setGeneric("get_accession_fdr", 
            function(.Object, ...) standardGeneric("get_accession_fdr"))
@@ -31,23 +33,50 @@ setGeneric("get_psm_fdr",
            function(.Object, ...) standardGeneric("get_psm_fdr"))
 
 setGeneric("get_accessions", 
-           function(.Object, ...) standardGeneric("get_accessions"))
+           function(.Object) standardGeneric("get_accessions"))
 
 setGeneric("get_peptides", 
-          function(.Object, ...) standardGeneric("get_peptides"))
+          function(.Object) standardGeneric("get_peptides"))
 
 
 setGeneric("read_mzIDs", 
-           function(.Object, ...) standardGeneric("read_mzIDs"))
+           function(.Object, mzids) standardGeneric("read_mzIDs"))
 
 
 setGeneric("apply_filter", 
-           function(.Object, .Filter, ...) standardGeneric("apply_filter"))
+           function(.Object, .Filter) standardGeneric("apply_filter"))
 
 
 setGeneric("evaluate_filter", 
[…]

### 2014-04-01 R/MSnID-methods.R

@@ -44,7 +44,7 @@
 
 
 .assess_missed_cleavages <- function(peptide, 
-                                    missedCleavagePattern="[KR](?=[^P$])")
+                                    missedCleavagePattern)
 {
    # the fastest approach
    peptide <- .get_clean_peptide_sequence(peptide)
@@ -57,11 +57,12 @@
 
 
 setMethod("assess_missed_cleavages", signature("MSnID"),
-          definition=function(.Object, ...)
+          definition=function(.Object, missedCleavagePattern)
           {
              .check_column(.Object, "Peptide")
              .Object@psms$NumMissCleavages <-
-                .assess_missed_cleavages(as.character(.Object@psms$Peptide), ...)
+                .assess_missed_cleavages(as.character(.Object@psms$Peptide), 
+                                         missedCleavagePattern)
              return(.Object)
           }
 )
@@ -69,7 +70,7 @@ setMethod("assess_missed_cleavages", signature("MSnID"),
 #---
 
 .assess_termini <- function(peptide, 
-                           validCleavagePattern="[RK]\\.[^P]")
+                           validCleavagePattern)
 {
    # "[RK]\\.[^P]" | "-\\."
    # It has to be sequence with flanking AAs e.g. K.XXXXXXXR.X
@@ -95,10 +96,11 @@ setMethod("assess_missed_cleavages", signature("MSnID"),
              
 
 setMethod("assess_termini", signature("MSnID"),
-          definition=function(.Object, ...)
+          definition=function(.Object, validCleavagePattern)
           {
              .Object@psms$NumIrregCleavages <- 
-                .assess_termini(as.character(.Object@psms$Peptide), ...)
+                .assess_termini(as.character(.Object@psms$Peptide), 
+                                validCleavagePattern)
              return(.Object)
           }
 )
@@ -327,13 +329,9 @@ MSnID <- function(workDir='.', cleanCache=FALSE)
                     "chargeState",
                     "spectrumFile", "spectrumID")
 
-setGeneric("psms",
-           function(.Object, ...) standardGeneric("psms"))
 setMethod("psms","MSnID", 
           function(.Object) as.data.frame(.Object@psms))
 
-setGeneric("psms<-",
-           function(.Object, value) standardGeneric("psms<-"))
 setReplaceMethod("psms",
                  signature(.Object="MSnID",
                            value="data.frame"),
@@ -355,8 +353,8 @@ setReplaceMethod("psms",
                  })
 
 
[…]

### 2014-04-01 R/MSnIDFilter-methods.R

@@ -2,32 +2,25 @@
 
 
 
-setGeneric("get_starting_parameters",
-           function(x, y, ...) standardGeneric("get_starting_parameters"))
+# setGeneric("get_starting_parameters",
+#            function(x, y, ...) standardGeneric("get_starting_parameters"))
+# 
+# setMethod("get_starting_parameters",
+#           signature(x="MSnIDFilter", y="MSnID"),
+#           definition=function(x, y, ...)
+#           {
+#              # let's assume we are taking median
+#              # update the filter list
+#              for( name_i in names(x@filterList)){
+#                 x@filterList[[name_i]]$threshold <- 
+#                    median(y[[name_i]],na.rm=TRUE)
+#              }
+#              return(x)
+#           }
+# )
 
 
 
-setMethod("get_starting_parameters",
-          signature(x="MSnIDFilter", y="MSnID"),
-          definition=function(x, y, ...)
-          {
-             # let's assume we are taking median
-             # update the filter list
-             for( name_i in names(x@filterList)){
-                x@filterList[[name_i]]$threshold <- 
-                   median(y[[name_i]],na.rm=TRUE)
-             }
-             return(x)
-          }
-)
-
-
-
-
-# # how about rule:
-# # if not exist - append,
-# # if exist - update threshold (leave comparison if exists)
-# # reset is separate function
 
 setMethod("set_filter", 
           signature(.Object="MSnIDFilter"), # dispatch just on the first argument
@@ -82,13 +75,13 @@ setAs("MSnIDFilter", "character",
 setAs("MSnIDFilter", "numeric",
       def=function(from) .get_filterValues(from, precision=Inf))
 
-# getGeneric("as.numeric")
 setMethod("as.numeric", "MSnIDFilter",
           definition=function(x, ...)
              as(x,"numeric"))
 
-setMethod("as.vector", "MSnIDFilter",
-          definition=function(x) as(x,"numeric"))
+# setMethod("as.vector", "MSnIDFilter",
+#           definition=function(x) 
+#              as(x,"numeric"))
 
 setMethod("length", "MSnIDFilter",
           definition=function(x) length(x@filterList))
@@ -101,15 +94,11 @@ setMethod("update", "MSnIDFilter",
           {
              #
              stopifnot(length(object) == length(newThresholds))
-             for(i in 1:length(object))
+             for(i in seq_along(object))
[…]

### 2014-04-01 R/optimize_filter.r

@@ -58,12 +58,6 @@
    newFilter <- update(filterObj, as.numeric(optim.pars))
    return(newFilter)
 }
-
-
-
-
-setGeneric("optimize_filter",
-           function(.Filter, .Data, ...) stadardGeneric("optimize_filter"))
            
            
 setMethod("optimize_filter",

### 2014-04-01 R/zzz.R

@@ -1,4 +1,4 @@
 
 .onLoad <- function(libname, pkgname) {
-   print("Starting MSnID...")
+#    packageStartupMessage("Starting MSnID...")
 }

### 2014-04-01 examples/c_elegans.R
deleting examples directory

@@ -1,231 +1,0 @@
-# --- DOWNLOADING MS/MS SEARCH RESULTS ---
-try(setInternet2(FALSE),silent=TRUE)
-ftp <- "ftp://PASS00308:PJ5348t@ftp.peptideatlas.org/"
-meta <- read.delim(sprintf("%sSample_Metadata.txt", ftp), as.is=T)
-meta <- subset(meta, Age == 'young' & Diet == 'ff') # explore the effect of daf-16
-for( dataset in meta$PNNL.Dataset.Name){
-   cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
-                       ftp, dataset)
-   download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
-}
-#----------------------------------------
-
-
-#--- READING MZID FILES ----------------
-library("MSnID")
-# start the project by providing work directory
-msnid <- MSnID(".")
-mzids <- list.files(".", pattern=".mzid.gz")
-msnid <- read_mzIDs(msnid, mzids)
-#---------------------------------------
-
-
-
-
-
-
-#--- CHECKING WHAT IS INSIDE -----------
-head(psms(msnid))
-names(msnid)
-show(msnid) # Key columns are not set
-#---------------------------------------
-
-
-
-
-# --- UPDATES TO INCLUDE KEY COLUMNS ---
-# Note, the anticipated/suggested columns in the
-# peptide-to-spectrum matching results are:
-#    Peptide, Accession, isDecoy, 
-#    calculatedMassToCharge, experimentalMassToCharge, 
-#    spectrumFile, spectrumID
-msnid$Peptide <- paste(msnid$pre, msnid$pepseq, msnid$post, sep='.')
-msnid$Accession <- msnid$accession
-msnid$isDecoy <- msnid$isdecoy
-msnid$calculatedMassToCharge <- msnid$calculatedmasstocharge
-msnid$experimentalMassToCharge <- msnid$experimentalmasstocharge
-msnid$chargeState <- msnid$chargestate
-msnid$spectrumID <- msnid$spectrumid
-# tidying up. removing left-over columns
-msnid$accession <- NULL
-msnid$pepSeq <- NULL
-msnid$pre <- NULL
-msnid$post <- NULL
-msnid$idecoy <- NULL
-msnid$spectrumid <- NULL
-msnid$experimentalmasstocharge <- NULL
[…]

### 2014-04-01 inst/doc/msnid_vignette.Rnw
adding inst/doc directory with vignette

--- 2014-03-28 vignette/msnid_vignette.Rnw
@@
 %\VignetteIndexEntry{MSnID Package for Handling MS/MS Identifications}
-% \VignetteDepends{BiocStyle, msmsTests}
+%\VignetteDepends{BiocStyle, msmsTests, ggplot2}
 %\VignetteKeywords{Documentation}
@@
 ftp <- "ftp://PASS00308:PJ5348t@ftp.peptideatlas.org/"
-meta <- read.delim(sprintf("%sSample_Metadata.txt", ftp), as.is=T)
+meta <- read.delim(sprintf("%sSample_Metadata.txt", ftp), as.is=TRUE)
 meta <- subset(meta, Age == 'young' & Diet == 'ff') # explore the effect of daf-16.
-#for( dataset in meta$PNNL.Dataset.Name){
-#   cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
-#                       ftp, dataset)
-#   download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
-#}
+for( dataset in meta$PNNL.Dataset.Name){
+   cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
+                       ftp, dataset)
+   download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
+}
 @

### 2014-04-01 demo/c_elegans.R
adding demo directory

@@ -1,0 +1,231 @@
+# --- DOWNLOADING MS/MS SEARCH RESULTS ---
+try(setInternet2(FALSE),silent=TRUE)
+ftp <- "ftp://PASS00308:PJ5348t@ftp.peptideatlas.org/"
+meta <- read.delim(sprintf("%sSample_Metadata.txt", ftp), as.is=TRUE)
+meta <- subset(meta, Age == 'young' & Diet == 'ff') # explore the effect of daf-16
+for( dataset in meta$PNNL.Dataset.Name){
+   cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
+                       ftp, dataset)
+   download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
+}
+#----------------------------------------
+
+
+#--- READING MZID FILES ----------------
+library("MSnID")
+# start the project by providing work directory
+msnid <- MSnID(".")
+mzids <- list.files(".", pattern=".mzid.gz")
+msnid <- read_mzIDs(msnid, mzids)
+#---------------------------------------
+
+
+
+
+
+
+#--- CHECKING WHAT IS INSIDE -----------
+head(psms(msnid))
+names(msnid)
+show(msnid) # Key columns are not set
+#---------------------------------------
+
+
+
+
+# --- UPDATES TO INCLUDE KEY COLUMNS ---
+# Note, the anticipated/suggested columns in the
+# peptide-to-spectrum matching results are:
+#    Peptide, Accession, isDecoy, 
+#    calculatedMassToCharge, experimentalMassToCharge, 
+#    spectrumFile, spectrumID
+msnid$Peptide <- paste(msnid$pre, msnid$pepseq, msnid$post, sep='.')
+msnid$Accession <- msnid$accession
+msnid$isDecoy <- msnid$isdecoy
+msnid$calculatedMassToCharge <- msnid$calculatedmasstocharge
+msnid$experimentalMassToCharge <- msnid$experimentalmasstocharge
+msnid$chargeState <- msnid$chargestate
+msnid$spectrumID <- msnid$spectrumid
+# tidying up. removing left-over columns
+msnid$accession <- NULL
+msnid$pepSeq <- NULL
+msnid$pre <- NULL
+msnid$post <- NULL
+msnid$idecoy <- NULL
+msnid$spectrumid <- NULL
+msnid$experimentalmasstocharge <- NULL
[…]

### 2014-04-02 demo/c_elegans.R
polished c_elegans example

@@ -3,11 +3,11 @@ try(setInternet2(FALSE),silent=TRUE)
 ftp <- "ftp://PASS00308:PJ5348t@ftp.peptideatlas.org/"
 meta <- read.delim(sprintf("%sSample_Metadata.txt", ftp), as.is=TRUE)
 meta <- subset(meta, Age == 'young' & Diet == 'ff') # explore the effect of daf-16
-for( dataset in meta$PNNL.Dataset.Name){
-   cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
-                       ftp, dataset)
-   download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
-}
+# for( dataset in meta$PNNL.Dataset.Name){
+#    cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
+#                        ftp, dataset)
+#    download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
+# }
 #----------------------------------------
 
 
@@ -20,17 +20,11 @@ msnid <- read_mzIDs(msnid, mzids)
 #---------------------------------------
 
 
-
-
-
-
 #--- CHECKING WHAT IS INSIDE -----------
 head(psms(msnid))
 names(msnid)
-show(msnid) # Key columns are not set
+show(msnid) # Key columns are not set yet.
 #---------------------------------------
-
-
 
 
 # --- UPDATES TO INCLUDE KEY COLUMNS ---
@@ -56,25 +50,35 @@ msnid$spectrumid <- NULL
 msnid$experimentalmasstocharge <- NULL
 msnid$calculatedmasstocharge <- NULL
 msnid$chargestate <- NULL
-#-------------------------------------
 show(msnid)
+#-------------------------------------
 
 
 # --- EXTRA INFO ABOUT PEPTIDE SEQUENCES ---
 msnid <- assess_termini(msnid, validCleavagePattern="[RK]\\.[^P]")
 msnid <- assess_missed_cleavages(msnid, missedCleavagePattern="[KR](?=[^P$])")
 # visualize distributions of missed cleavages as an example
-library("ggplot2")
 pepCleav <- unique(psms(msnid)[,c("NumMissCleavages", "isDecoy", "Peptide")])
 pepCleav <- as.data.frame(table(pepCleav[,c("NumMissCleavages", "isDecoy")]))
[…]

### 2014-04-02 inst/doc/msnid_vignette.Rnw
tweaked vignette

@@
 x <- data.frame(dM, isDecoy=msnid$isDecoy)
-ggplot(x, aes(x=dM, fill=isDecoy)) + geom_histogram(position='stack')
+ggplot(x, aes(x=dM, fill=isDecoy)) + 
+   geom_histogram(position='stack', binwidth=0.1)
 @
@@
 filtObj$absParentMassErrorPPM <- list(comparison="<", threshold=10.0)
-filtObj$msmsScore <- list(comparison=">", threshold=7.0)
+filtObj$msmsScore <- list(comparison=">", threshold=10.0)
 # print filter
@@
 
-# Nelder-Mead optimization
-set.seed(0)
+# Fine tuning. Nelder-Mead optimization
 system.time({
@@
 # (absParentMassErrorPPM < 3) & (msmsScore > 7.8)
-
-# simulated annealing optimization
-set.seed(0)
-system.time({
-   filtObj.sann <- optimize_filter(filtObj.grid, msnid, fdr.max=0.01, 
-                                  method="SANN", level="Peptide", n.iter=500)})
-show(filtObj.sann)
-# (absParentMassErrorPPM < 2.2) & (msmsScore > 7.6)
 @
-Let's check which filter turned out to be the best. The differece between the optimized filters is likely to be very small.
+Evaluating filter based on a good guess and optimization. Finally we'll apply
+optimized filter to proceed with further steps in the pipeline.
 <<>>=
+# comparing orignal and optimized filters
 evaluate_filter(msnid, filtObj, level="Peptide")
-evaluate_filter(msnid, filtObj.grid, level="Peptide")
 evaluate_filter(msnid, filtObj.nm, level="Peptide")
-evaluate_filter(msnid, filtObj.sann, level="Peptide")
+# apply the optimized filter
+msnid <- apply_filter(msnid, filtObj.nm)
+show(msnid)
 @
-We'll choose and apply filter produced by simulated annealing optimization procedure.
 Identifications that matched to decoy and contaminant protein sequences can be removed by providing filters in the forms of text strings that will be evaluated in the context of PSM table.
 <<>>=
-# apply the optimized filter
-msnid <- apply_filter(msnid, filtObj.sann)
-show(msnid)
 # let's remove reverse/decoy and Contaminants
@@
 \end{center}
-\pagebreak[4]
+\pagebreak[0]
 <<label=rollingToProteinLevel>>==

### 2014-04-02 demo/c_elegans.R
Update c_elegans.R
remove left-over comments that skip download of mzIdentMLs

@@ -3,11 +3,11 @@ try(setInternet2(FALSE),silent=TRUE)
 ftp <- "ftp://PASS00308:PJ5348t@ftp.peptideatlas.org/"
 meta <- read.delim(sprintf("%sSample_Metadata.txt", ftp), as.is=TRUE)
 meta <- subset(meta, Age == 'young' & Diet == 'ff') # explore the effect of daf-16
-# for( dataset in meta$PNNL.Dataset.Name){
-#    cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
-#                        ftp, dataset)
-#    download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
-# }
+for( dataset in meta$PNNL.Dataset.Name){
+ cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
+                     ftp, dataset)
+ download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
+}
 #----------------------------------------
 
 

### 2014-04-02 demo/c_elegans.R
work on demo script

@@ -3,11 +3,11 @@ try(setInternet2(FALSE),silent=TRUE)
 ftp <- "ftp://PASS00308:PJ5348t@ftp.peptideatlas.org/"
 meta <- read.delim(sprintf("%sSample_Metadata.txt", ftp), as.is=TRUE)
 meta <- subset(meta, Age == 'young' & Diet == 'ff') # explore the effect of daf-16
-# for( dataset in meta$PNNL.Dataset.Name){
-#    cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
-#                        ftp, dataset)
-#    download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
-# }
+for( dataset in meta$PNNL.Dataset.Name){
+   cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
+                       ftp, dataset)
+   download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
+}
 #----------------------------------------
 
 
@@ -252,5 +252,13 @@ lst <- test.results(res,msnset.prot,pData(msnset.prot)$Daf.16.type,"wt","mut",di
                     method="BH")
 res.volcanoplot(lst$tres, min.LFC=1, max.pval=0.05, ylbls=NULL, maxy=4)
 #--------------------------------------
+
+
+
+regulated <- subset(lst$tres, adjp < 0.05 & abs(LogFC) > 1)
+selected.data <- exprs(msnset.prot[rownames(regulated),])
+library("gplots")
+heatmap.2(selected.data)
+library("Heatplus")
 
 

### 2014-04-02 demo/c_elegans.R
.

@@ -4,15 +4,9 @@ ftp <- "ftp://PASS00308:PJ5348t@ftp.peptideatlas.org/"
 meta <- read.delim(sprintf("%sSample_Metadata.txt", ftp), as.is=TRUE)
 meta <- subset(meta, Age == 'young' & Diet == 'ff') # explore the effect of daf-16
 for( dataset in meta$PNNL.Dataset.Name){
-<<<<<<< HEAD
    cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
                        ftp, dataset)
    download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
-=======
- cel.path <- sprintf("%s/MSGFPlus_Results/MZID_Files/%s_msgfplus.mzid.gz", 
-                     ftp, dataset)
- download.file(cel.path, sprintf("%s_msgfplus.mzid.gz", dataset))
->>>>>>> 0e9d43c18841ea5ac87beb4af955441d64aa13ee
 }
 #----------------------------------------
 
@@ -260,7 +254,7 @@ res.volcanoplot(lst$tres, min.LFC=1, max.pval=0.05, ylbls=NULL, maxy=4)
 #--------------------------------------
 
 
-
+# --- work in progress ---
 regulated <- subset(lst$tres, adjp < 0.05 & abs(LogFC) > 1)
 selected.data <- exprs(msnset.prot[rownames(regulated),])
 library("gplots")

### 2014-04-07 demo/c_elegans.R
added heatmap

@@ -254,11 +254,25 @@ res.volcanoplot(lst$tres, min.LFC=1, max.pval=0.05, ylbls=NULL, maxy=4)
 #--------------------------------------
 
 
-# --- work in progress ---
-regulated <- subset(lst$tres, adjp < 0.05 & abs(LogFC) > 1)
-selected.data <- exprs(msnset.prot[rownames(regulated),])
-library("gplots")
-heatmap.2(selected.data)
-library("Heatplus")
 
+# --- HEATMAP ---
+if(!require("Heatplus")){
+   library("BiocInstaller")
+   biocLite("Heatplus")
+   library("Heatplus")
+}
+regulated <- subset(lst$tres, adjp < 0.05 & abs(LogFC) > 1)
+# order MSnSet object the daf-16 status
+msnset <- msnset[,order(pData(msnset)$Daf.16.type)]
+# matrix with regulated proteins
+selected.data <- exprs(msnset[rownames(regulated),])
+# more meaningful sample names
+colnames(selected.data) <- with(pData(msnset), 
+                                paste(Daf.16.type, Letter.Replicate, sep='.'))
+# scaling counts from 0 to 1
+selected.data <- sweep(selected.data, 1, apply(selected.data, 1, min), '-')
+selected.data <- sweep(selected.data, 1, apply(selected.data, 1, max), '/')
+heatmap_plus(selected.data,
+             scale='none',
+             col=colorRampPalette(c("snow","steelblue"))(10))
 

### 2014-04-07 demo/c_elegans.R
subsetting param data.frame to accelerate plotting

@@ -134,6 +134,9 @@ ggplot(params) +
    geom_density(aes(x = msmsScore, color = isDecoy, ..count..))
 ggplot(params) + 
    geom_density(aes(x = absParentMassErrorPPM, color = isDecoy, ..count..))
+# subsetting params to top 10000 to accelerate the plotting
+set.seed(0)
+params <- params[sample.int(nrow(msnid),10000),]
 ggplot(data=params, aes(x=msmsScore, y=absParentMassErrorPPM, color=isDecoy)) +
    geom_point(size=1.5)
 # Setting up filter object

### 2014-04-07 demo/c_elegans.R
fix msnset object name
msnset to msnset.prot

@@ -266,11 +266,11 @@ if(!require("Heatplus")){
 }
 regulated <- subset(lst$tres, adjp < 0.05 & abs(LogFC) > 1)
 # order MSnSet object the daf-16 status
-msnset <- msnset[,order(pData(msnset)$Daf.16.type)]
+msnset.prot <- msnset.prot[,order(pData(msnset.prot)$Daf.16.type)]
 # matrix with regulated proteins
-selected.data <- exprs(msnset[rownames(regulated),])
+selected.data <- exprs(msnset.prot[rownames(regulated),])
 # more meaningful sample names
-colnames(selected.data) <- with(pData(msnset), 
+colnames(selected.data) <- with(pData(msnset.prot), 
                                 paste(Daf.16.type, Letter.Replicate, sep='.'))
 # scaling counts from 0 to 1
 selected.data <- sweep(selected.data, 1, apply(selected.data, 1, min), '-')
@@ -278,4 +278,3 @@ selected.data <- sweep(selected.data, 1, apply(selected.data, 1, max), '/')
 heatmap_plus(selected.data,
              scale='none',
              col=colorRampPalette(c("snow","steelblue"))(10))
-

### 2014-04-07 R/MSnIDFilter-methods.R
fix typo in show

@@ -1,7 +1,6 @@
 
 
 
-
 # setGeneric("get_starting_parameters",
 #            function(x, y, ...) standardGeneric("get_starting_parameters"))
 # 
@@ -103,8 +102,8 @@ setMethod("update", "MSnIDFilter",
 setMethod("show", "MSnIDFilter",
           definition=function(object)
           {
-             cat("An object of calss \"",class(object),"\"\n",sep='')
-             cat("Filter as string:\n")
+             cat("An object of class \"",class(object),"\"\n",sep='')
+             #cat("Filter as string:\n")
              cat(.get_filterString(object),'\n')
              #cat(capture.output(str(filterList))[-1],sep='\n')
              

### 2014-04-07 R/MSnIDFilter-methods.R
tweak show

@@ -1,6 +1,5 @@
 
 
-
 # setGeneric("get_starting_parameters",
 #            function(x, y, ...) standardGeneric("get_starting_parameters"))
 # 
@@ -102,7 +101,8 @@ setMethod("update", "MSnIDFilter",
 setMethod("show", "MSnIDFilter",
           definition=function(object)
           {
-             cat("An object of class \"",class(object),"\"\n",sep='')
+             #cat("An object of class \"",class(object),"\"\n",sep='')
+             cat(class(object)," object\n",sep='')
              #cat("Filter as string:\n")
              cat(.get_filterString(object),'\n')
              #cat(capture.output(str(filterList))[-1],sep='\n')

### 2014-04-08 inst/doc/msnid_vignette.Rnw
wholesale commit

@@
 # creates new column NumIrregCleavages
-msnid <- assess_termini(msnid, validCleavagePattern="[RK]\\.[^P]")
+msnid <- assess_termini(msnid, validCleavagePattern="[KR]\\.[^P]")
 # creates new column NumMissCleavages

### 2014-04-08 R/MSnIDFilter-methods.R

@@ -103,7 +103,7 @@ setMethod("update", "MSnIDFilter",
 setMethod("show", "MSnIDFilter",
           definition=function(object)
           {
-             cat("An object of calss \"",class(object),"\"\n",sep='')
+             cat("An object of class \"",class(object),"\"\n",sep='')
              cat("Filter as string:\n")
              cat(.get_filterString(object),'\n')
              #cat(capture.output(str(filterList))[-1],sep='\n')

### 2014-04-08 demo/c_elegans.R

@@ -185,6 +185,12 @@ show(msnid)
 #---------------------------------------------
 
 
+# msnid$msgfpepqvalue <- msnid$`ms-gf:pepqvalue`
+# ff <- MSnIDFilter(msnid)
+# ff$msgfpepqvalue <- list(comparison="<", threshold=0.01)
+# evaluate_filter(msnid, ff, level="Peptide")
+
+
 # --- CONVERTING TO MSnSet -------------------
 msnset <- as(msnid, "MSnSet")
 # Note, feature data is peptide-centric. Peptide to protein assigments
@@ -255,10 +261,32 @@ res.volcanoplot(lst$tres, min.LFC=1, max.pval=0.05, ylbls=NULL, maxy=4)
 
 
 # --- work in progress ---
+# selecting regulated only
 regulated <- subset(lst$tres, adjp < 0.05 & abs(LogFC) > 1)
+# order MSnSet object the daf-16 status
+msnset.prot <- msnset.prot[,order(pData(msnset.prot)$Daf.16.type)]
+# matrix with regulated proteins
 selected.data <- exprs(msnset.prot[rownames(regulated),])
+# more meaningful sample names
+colnames(selected.data) <- with(pData(msnset.prot), 
+                                paste(Daf.16.type, Letter.Replicate, sep='.'))
+# scaling counts from 0 to 1
+selected.data <- sweep(selected.data, 1, apply(selected.data, 1, min), '-')
+selected.data <- sweep(selected.data, 1, apply(selected.data, 1, max), '/')
+group.colors <- c('green','red')[as.factor(pData(msnset.prot)$Daf.16.type)]
 library("gplots")
-heatmap.2(selected.data)
-library("Heatplus")
+heatmap.2(selected.data,
+          Colv=FALSE,
+          dendrogram='row',
+          col=colorRampPalette(c("snow","steelblue"))(10),
+          ColSideColors=group.colors,
+          key=FALSE,
+          trace='none',
+          main='Scaled Spectral Counts for ')
+
+
+
+
+
 
 

### 2014-04-08 demo/c_elegans.R
fix demo

@@ -188,10 +188,6 @@ show(msnid)
 #---------------------------------------------
 
 
-# msnid$msgfpepqvalue <- msnid$`ms-gf:pepqvalue`
-# ff <- MSnIDFilter(msnid)
-# ff$msgfpepqvalue <- list(comparison="<", threshold=0.01)
-# evaluate_filter(msnid, ff, level="Peptide")
 
 
 # --- CONVERTING TO MSnSet -------------------
@@ -263,18 +259,12 @@ res.volcanoplot(lst$tres, min.LFC=1, max.pval=0.05, ylbls=NULL, maxy=4)
 #--------------------------------------
 
 
-<<<<<<< HEAD
-# --- work in progress ---
-# selecting regulated only
-=======
-
 # --- HEATMAP ---
 if(!require("Heatplus")){
    library("BiocInstaller")
    biocLite("Heatplus")
    library("Heatplus")
 }
->>>>>>> 96dc4166f307d747cdb7d9d9d4b2867c4cd40b45
 regulated <- subset(lst$tres, adjp < 0.05 & abs(LogFC) > 1)
 # order MSnSet object the daf-16 status
 msnset.prot <- msnset.prot[,order(pData(msnset.prot)$Daf.16.type)]
@@ -286,26 +276,6 @@ colnames(selected.data) <- with(pData(msnset.prot),
 # scaling counts from 0 to 1
 selected.data <- sweep(selected.data, 1, apply(selected.data, 1, min), '-')
 selected.data <- sweep(selected.data, 1, apply(selected.data, 1, max), '/')
-<<<<<<< HEAD
-group.colors <- c('green','red')[as.factor(pData(msnset.prot)$Daf.16.type)]
-library("gplots")
-heatmap.2(selected.data,
-          Colv=FALSE,
-          dendrogram='row',
-          col=colorRampPalette(c("snow","steelblue"))(10),
-          ColSideColors=group.colors,
-          key=FALSE,
-          trace='none',
-          main='Scaled Spectral Counts for ')
-
-
-
-
-
-
-
-=======
 heatmap_plus(selected.data,
              scale='none',
              col=colorRampPalette(c("snow","steelblue"))(10))
->>>>>>> 96dc4166f307d747cdb7d9d9d4b2867c4cd40b45

### 2014-07-11 R/MSnID-methods.R
change to data.table style handling

@@ -387,10 +387,11 @@ setMethod("read_mzIDs", "MSnID",
 
 
 
+#-------------------------------------------------------------------------------
 # Getters and setters modelled after
 # library(Biobase)
-# getMethod("$", "eSet")
-# getMethod("$<-", "eSet")
+# getMethod("$", "eSet") # tricky
+# getMethod("$<-", "eSet") # straight
 # getMethod("[[", "eSet")
 # getMethod("[[<-", "eSet")
 
@@ -398,7 +399,7 @@ setMethod("$", "MSnID",
           definition=function(x, name)
           {
              eval(substitute(x@psms$NAME_ARG, list(NAME_ARG = name)))
-             # return(x@psms[[name]])
+             return(x@psms[[name]])
           }
 )
 
@@ -409,13 +410,32 @@ setMethod("[[", "MSnID",
           }
 )
 
+
+
+# # data.frame style
+# setMethod("$<-", "MSnID",
+#           definition=function(x, name, value)
+#           {
+#              x@psms[[name]] <- value
+#              return(x)
+#           }
+# )
+
+# data.table style
+# data(c_elegans)
+# debug(getMethod("$<-", "MSnID"))
+# msnidObj$HREN <- 1
 setMethod("$<-", "MSnID",
           definition=function(x, name, value)
           {
-             x@psms[[name]] <- value
+             x@psms <- eval(substitute(x@psms[,NAME_ARG:=value], list(NAME_ARG=name)))
              return(x)
           }
 )
+# to debug
+# trace("$<-", browser, exit=browser, signature = c("MSnID"))
+
+
 
 setMethod("[[<-", "MSnID",
           definition=function(x, i, j, ..., value)
@@ -424,6 +444,9 @@ setMethod("[[<-", "MSnID",
              return(x)
           }
 )
+#-------------------------------------------------------------------------------
+
+
 
 
 

### 2014-07-11 R/read_mzIDs.R
band-aid for letter case handling

@@ -10,7 +10,64 @@
       data <- data.table(flatten(mzID(mzids)))
       saveCache(data, key=key)
    }
+   
+   # the piece below will be removed in the future,
+   # one the corresponding functionality implemented in mzID
+   # do some cleaning/tweaking of column modes and name casing
+   
+   # Columns that must be present:
+   # Peptide, Accession, isDecoy, calculatedMassToCharge, 
+   # experimentalMassToCharge, chargeState, spectrumFile, spectrumID
+   if(!is.null(data$accession)){
+      data$Accession <- data$accession
+      data$accession <- NULL
+   }
+
+   if(!is.null(data$isdecoy)){
+      data$isDecoy <- data$isdecoy
+      data$isdecoy <- NULL
+   }
+   
+   if(!is.null(data$calculatedmasstocharge)){
+      data$calculatedMassToCharge <- as.numeric(data$calculatedmasstocharge)
+      data$calculatedmasstocharge <- NULL
+   }
+   
+   if(!is.null(data$experimentalmasstocharge)){
+      data$experimentalMassToCharge <- as.numeric(data$experimentalmasstocharge)
+      data$experimentalmasstocharge <- NULL
+   }
+   
+   if(!is.null(data$chargestate)){
+      data$chargeState <- data$chargestate
+      data$chargestate <- NULL
+   }
+   
+   if(!is.null(data$spectrumid)){
+      data$spectrumID <- data$spectrumid
+      data$spectrumid <- NULL
+   }
+   
+   if(!is.null(data$pre)){
+      data$Pre <- data$pre
+      data$pre <- NULL
+   }
+   
+   if(!is.null(data$post)){
+      data$Post <- data$post
+      data$post <- NULL
+   }
+   
+   if(!is.null(data$pepseq)){
+      data$PepSeq <- data$pepseq
+      data$pepseq <- NULL
+   }
+   #---
+   data$Peptide <- paste(data$Pre, data$PepSeq, data$Post, sep='.')
+   
    return(data)
 }
+
+
 
 

### 2014-07-11 inst/unitTests/test_data_load.R
adjusting to name handling in read_mzIDs

@@ -2,22 +2,22 @@ library("MSnID")
 msnid2 <- MSnID(".")
 mzids <- system.file("extdata","c_elegans.mzid.gz",package="MSnID")
 msnid2 <- read_mzIDs(msnid2, mzids)
-msnid2$Accession <- msnid2$accession
-msnid2$accession <- NULL
-msnid2$isDecoy <- msnid2$isdecoy
-msnid2$isdecoy <- NULL
-msnid2$calculatedMassToCharge <- msnid2$calculatedmasstocharge
-msnid2$calculatedmasstocharge <- NULL
-msnid2$experimentalMassToCharge <- msnid2$experimentalmasstocharge
-msnid2$experimentalmasstocharge <- NULL
-msnid2$chargeState <- msnid2$chargestate
-msnid2$chargestate <- NULL
-msnid2$spectrumID <- msnid2$spectrumid
-msnid2$spectrumid <- NULL
-msnid2$Peptide <- paste(msnid2$pre, msnid2$pepseq, msnid2$post, sep='.')
-msnid2$pre <- NULL
-msnid2$post <- NULL
-msnid2$pepseq <- NULL
+# msnid2$Accession <- msnid2$accession
+# msnid2$accession <- NULL
+# msnid2$isDecoy <- msnid2$isdecoy
+# msnid2$isdecoy <- NULL
+# msnid2$calculatedMassToCharge <- msnid2$calculatedmasstocharge
+# msnid2$calculatedmasstocharge <- NULL
+# msnid2$experimentalMassToCharge <- msnid2$experimentalmasstocharge
+# msnid2$experimentalmasstocharge <- NULL
+# msnid2$chargeState <- msnid2$chargestate
+# msnid2$chargestate <- NULL
+# msnid2$spectrumID <- msnid2$spectrumid
+# msnid2$spectrumid <- NULL
+# msnid2$Peptide <- paste(msnid2$Pre, msnid2$PepSeq, msnid2$Post, sep='.')
+# msnid2$pre <- NULL
+# msnid2$post <- NULL
+# msnid2$pepseq <- NULL
 # clean up the cache directory
 unlink(".Rcache", recursive=TRUE)
 

### 2014-08-05 R/zzz.R
no need for start-up message

@@ -1,4 +1,0 @@
-
-.onLoad <- function(libname, pkgname) {
-#    packageStartupMessage("Starting MSnID...")
-}

### 2014-08-05 R/AllClasses.R
changing tabs

@@ -3,20 +3,18 @@
 .PROTON_MASS <- 1.007276466812
 
 setClass(Class="MSnID",
-         representation(
-            workDir="character", # working directory
-            psms="data.table" # peptide-to-spectra matches
-            )
-         )
+            representation(
+                workDir="character", # working directory
+                psms="data.table") # peptide-to-spectra matches
+)
 
 
 
 
 # partitions and parameter transforms to be added to filter class
 setClass(Class="MSnIDFilter",
-         representation(
-            filterList="list",
-            validParNames="character"
-         )
+            representation(
+                filterList="list",
+                validParNames="character")
 )
 

### 2014-08-05 R/AllGenerics.R
changing tabs from 3 to 4

@@ -1,68 +1,72 @@
 
 setGeneric("set_filter",
-           function(.Object, ...) standardGeneric("set_filter"))
+            function(.Object, ...) standardGeneric("set_filter"))
 
 setGeneric("assess_missed_cleavages",
-           function(.Object, missedCleavagePattern="[KR](?=[^P$])") 
-              standardGeneric("assess_missed_cleavages"))
+            function(.Object, missedCleavagePattern="[KR](?=[^P$])") 
+                standardGeneric("assess_missed_cleavages"))
 
 # this may be an odd way to handle assess_termini
 setGeneric("assess_termini", 
-           function(.Object, validCleavagePattern="[KR]\\.[^P]") 
-              standardGeneric("assess_termini"))
+            function(.Object, validCleavagePattern="[KR]\\.[^P]") 
+                standardGeneric("assess_termini"))
 
 setGeneric("get_accession_fdr", 
-           function(.Object, ...) standardGeneric("get_accession_fdr"))
+            function(.Object, ...) standardGeneric("get_accession_fdr"))
 
 setGeneric("get_peptide_fdr", 
-           function(.Object, ...) standardGeneric("get_peptide_fdr"))
+            function(.Object, ...) standardGeneric("get_peptide_fdr"))
 
 setGeneric("get_psm_fdr", 
-           function(.Object, ...) standardGeneric("get_psm_fdr"))
+            function(.Object, ...) standardGeneric("get_psm_fdr"))
 
 setGeneric("get_accessions", 
-           function(.Object) standardGeneric("get_accessions"))
+            function(.Object) standardGeneric("get_accessions"))
 
 setGeneric("get_peptides", 
-          function(.Object) standardGeneric("get_peptides"))
+            function(.Object) standardGeneric("get_peptides"))
 
+
+# setGeneric("read_mzIDs", 
+#            function(.Object, mzids) standardGeneric("read_mzIDs"))
 
 setGeneric("read_mzIDs", 
-           function(.Object, mzids) standardGeneric("read_mzIDs"))
+            function(object, mzids) standardGeneric("read_mzIDs"))
+
 
 
 setGeneric("apply_filter", 
-           function(.Object, .Filter) standardGeneric("apply_filter"))
+            function(.Object, .Filter) standardGeneric("apply_filter"))
 
 
 setGeneric("evaluate_filter", 
-           function(.Object, filter, level=c("PSM", "Peptide", "Accession")) 
[…]

### 2014-08-05 R/MSnIDFilter-methods.R
tabs

@@ -1,23 +1,24 @@
 
 
 
-setMethod("set_filter", 
-          signature(.Object="MSnIDFilter"), # dispatch just on the first argument
-          definition=function(.Object, 
-                              parName,          # although this MUST be character
-                              comparison=NULL,
-                              threshold=NULL)
-          {
-             parNames <- sapply(.Object@filterList, "[[", "name")
-             parNames <- sapply(.Object@filterList, "[[", "name")
-          }
-)
+# setMethod("set_filter", 
+#             # dispatch just on the first argument
+#             signature(.Object="MSnIDFilter"), 
+#             definition=function(.Object, 
+#                                 parName,   # MUST be character
+#                                 comparison=NULL,
+#                                 threshold=NULL)
+#             {
+#                 parNames <- sapply(.Object@filterList, "[[", "name")
+#                 parNames <- sapply(.Object@filterList, "[[", "name")
+#             }
+# )
 
 
 .is_filterList_valid <- function(filterList)
 {
-   entries=c("comparison","threshold")
-   all(sapply(filterList, function(v) all(entries %in% names(v))))
+    entries=c("comparison","threshold")
+    all(sapply(filterList, function(v) all(entries %in% names(v))))
 }
 
 
@@ -25,149 +26,133 @@ setMethod("set_filter",
 # internal
 .get_filterString <- function(.Object, precision=2)
 {
-   # 1) assuming names in the list match names of list elements
-   # 2) assuming the names are in the rigth order
-   res <- lapply(.Object@filterList, 
-                 within, 
-                 {threshold <- signif(threshold, get("precision", parent.frame(n=4)))})
-   res <- lapply(names(res), 
-                 function(v) 
-                    paste(c(list(v), res[[v]]), collapse=' '))
-   res <- paste(sprintf("(%s)", res), collapse=" & ")
-   return(res)
+    # 1) assuming names in the list match names of list elements
+    # 2) assuming the names are in the rigth order
+    res <- lapply(.Object@filterList, 
+                    within, 
+                    {threshold <- signif(threshold, 
+                                        get("precision", parent.frame(n=4)))})
+    res <- lapply(names(res), 
+                    function(v) 
+                        paste(c(list(v), res[[v]]), collapse=' '))
+    res <- paste(sprintf("(%s)", res), collapse=" & ")
+    return(res)
 }
 
 
 .get_filterValues <- function(.Object, precision=2)
 {
[…]

### 2014-08-05 inst/unitTests/test_data_load.R
no need to update PSM column names, since it is handled by mzID package now

@@ -2,23 +2,6 @@ library("MSnID")
 msnid2 <- MSnID(".")
 mzids <- system.file("extdata","c_elegans.mzid.gz",package="MSnID")
 msnid2 <- read_mzIDs(msnid2, mzids)
-# msnid2$Accession <- msnid2$accession
-# msnid2$accession <- NULL
-# msnid2$isDecoy <- msnid2$isdecoy
-# msnid2$isdecoy <- NULL
-# msnid2$calculatedMassToCharge <- msnid2$calculatedmasstocharge
-# msnid2$calculatedmasstocharge <- NULL
-# msnid2$experimentalMassToCharge <- msnid2$experimentalmasstocharge
-# msnid2$experimentalmasstocharge <- NULL
-# msnid2$chargeState <- msnid2$chargestate
-# msnid2$chargestate <- NULL
-# msnid2$spectrumID <- msnid2$spectrumid
-# msnid2$spectrumid <- NULL
-# msnid2$Peptide <- paste(msnid2$Pre, msnid2$PepSeq, msnid2$Post, sep='.')
-# msnid2$pre <- NULL
-# msnid2$post <- NULL
-# msnid2$pepseq <- NULL
-# clean up the cache directory
 unlink(".Rcache", recursive=TRUE)
 
 
@@ -27,7 +10,7 @@ unlink(".Rcache", recursive=TRUE)
 # So I'll compare psms as data.frame only.
 
 test_data_load <- function() {
-   # now, check if it is what it supposed to be
-   data(c_elegans)
-   checkIdentical(psms(msnidObj), psms(msnid2))
+    # now, check if it is what it supposed to be
+    data(c_elegans)
+    checkIdentical(psms(msnidObj), psms(msnid2))
 }

### 2014-08-05 demo/c_elegans.R
changing tabs from 3 to 4 spaces

@@ -51,8 +51,8 @@ pepCleav <- unique(psms(msnid)[,c("NumMissCleavages", "isDecoy", "Peptide")])
 pepCleav <- as.data.frame(table(pepCleav[,c("NumMissCleavages", "isDecoy")]))
 library("ggplot2")
 ggplot(pepCleav, aes(x=NumMissCleavages, y=Freq, fill=isDecoy)) + 
-   geom_bar(stat="identity", position="dodge") + 
-   ggtitle("Number of Missed Cleavages")
+    geom_bar(stat="identity", position="dodge") + 
+    ggtitle("Number of Missed Cleavages")
 # number of cysteins per peptide sequence as an example of sequence analysis
 msnid$NumCys <- sapply(lapply(strsplit(msnid$Peptide,''),'==','C'),sum)
 # calculating peptide lengths
@@ -60,8 +60,8 @@ msnid$PepLength <- nchar(msnid$Peptide) - 4
 pepLen <- unique(psms(msnid)[,c("PepLength", "isDecoy", "Peptide")])
 # distribution of peptide lengths in the dataset
 ggplot(pepLen, aes(x=PepLength, fill=isDecoy)) + 
-   geom_histogram(position='dodge', binwidth=3) +
-   ggtitle("Distribution on of Peptide Lengths")
+    geom_histogram(position='dodge', binwidth=3) +
+    ggtitle("Distribution on of Peptide Lengths")
 #------------------------------------------
 
 
@@ -82,52 +82,53 @@ show(msnid)
 # original mass measurement error in ppm
 ppm <- mass_measurement_error(msnid)
 ggplot(as.data.frame(ppm), aes(x=ppm)) + 
-   geom_histogram(binwidth=100)
+    geom_histogram(binwidth=100)
 # In this particular case the problem that it goes
 # far is that there was no certainity that the reported
 # masses are the masses of monoisotopic peaks.
 # Let's take a look at mass measurement error in Dalton units.
-dM <- with(psms(msnid), (experimentalMassToCharge-calculatedMassToCharge)*chargeState)
+dM <- with(psms(msnid), 
+            (experimentalMassToCharge-calculatedMassToCharge)*chargeState)
 x <- data.frame(dM, isDecoy=msnid$isDecoy)
 ggplot(x, aes(x=dM, fill=isDecoy)) + 
-   geom_histogram(position='stack', binwidth=0.1)
+    geom_histogram(position='stack', binwidth=0.1)
 # Fixing the peak picking problem
[…]

### 2014-08-05 inst/unitTests/test_apply_filter.R

@@ -2,36 +2,36 @@ library("MSnID")
 data(c_elegans)
 
 test_apply_filter_1 <- function(){
-   # checking string filter
-   out <- apply_filter(msnidObj, "`ms-gf:specevalue` < 10^-10")
-   checkEquals(id_quality(out, level="PSM"),
-               list(fdr=0.000285103349964362, n=7017))
-   checkEquals(id_quality(out, level="Peptide"),
-               list(fdr=0.000778816199376947, n=2570))
-   checkEquals(id_quality(out, level="Accession"),
-               list(fdr=0.00193986420950533, n=1033))
+    # checking string filter
+    out <- apply_filter(msnidObj, "`MS-GF:SpecEValue` < 10^-10")
+    checkEquals(id_quality(out, level="PSM"),
+                list(fdr=0.000285103349964362, n=7017))
+    checkEquals(id_quality(out, level="peptide"),
+                list(fdr=0.000778816199376947, n=2570))
+    checkEquals(id_quality(out, level="accession"),
+                list(fdr=0.00193986420950533, n=1033))
 }
 
 
 test_apply_filter_2 <- function(){
-   # checking object-based filter
-   
-   msnidObj$msmsScore <- -log10(msnidObj$`ms-gf:specevalue`)
-   msnidObj$mzError <- abs(msnidObj$experimentalMassToCharge - 
-                              msnidObj$calculatedMassToCharge)
-   # setting up filter object
-   filtObj <- MSnIDFilter(msnidObj)
-   filtObj$msmsScore <- list(comparison=">", threshold=9.0)
-   filtObj$mzError <- list(comparison="<", threshold=0.2)
-   
-   out <- apply_filter(msnidObj, filtObj)
-   
-   checkEquals(id_quality(out, level="PSM"),
-               list(fdr=0.0011079650375477, n=8132))
-   checkEquals(id_quality(out, level="Peptide"),
-               list(fdr=0.0013550135501355, n=2956))
-   checkEquals(id_quality(out, level="Accession"),
-               list(fdr=0.00801424755120214, n=1132))
+    # checking object-based filter
+
+    msnidObj$msmsScore <- -log10(msnidObj$`MS-GF:SpecEValue`)
+    msnidObj$mzError <- abs(msnidObj$experimentalMassToCharge - 
+                            msnidObj$calculatedMassToCharge)
+    # setting up filter object
+    filtObj <- MSnIDFilter(msnidObj)
[…]

### 2014-08-05 R/read_mzIDs.R
No need to update column names. It is handled by mzID now.

@@ -1,71 +1,34 @@
 
-.read_mzIDs.memoized <- function(mzids) {
-   # Try to load cached data, if exists
-   key <- list(mzids)
-   data <- loadCache(key)
-   if (!is.null(data)) {
-      cat("Loaded cached data\n")
-   }else{   
-      cat("Reading from mzIdentMLs ...\n")
-      data <- data.table(flatten(mzID(mzids)))
-      saveCache(data, key=key)
-   }
-   
-   # the piece below will be removed in the future,
-   # one the corresponding functionality implemented in mzID
-   # do some cleaning/tweaking of column modes and name casing
-   
-   # Columns that must be present:
-   # Peptide, Accession, isDecoy, calculatedMassToCharge, 
-   # experimentalMassToCharge, chargeState, spectrumFile, spectrumID
-   if(!is.null(data$accession)){
-      data$Accession <- data$accession
-      data$accession <- NULL
-   }
+.read_mzIDs.memoized <- function(mzids)
+{
+    # Try to load cached data, if exists
+    key <- list(mzids)
+    data <- loadCache(key)
+    if (!is.null(data)){
+        cat("Loaded cached data\n")
+    }else{   
+        cat("Reading from mzIdentMLs ...\n")
+        data <- data.table(flatten(mzID(mzids), safeNames=FALSE))
+        saveCache(data, key=key)
+    }
 
-   if(!is.null(data$isdecoy)){
-      data$isDecoy <- data$isdecoy
-      data$isdecoy <- NULL
-   }
-   
-   if(!is.null(data$calculatedmasstocharge)){
-      data$calculatedMassToCharge <- as.numeric(data$calculatedmasstocharge)
-      data$calculatedmasstocharge <- NULL
-   }
-   
-   if(!is.null(data$experimentalmasstocharge)){
-      data$experimentalMassToCharge <- as.numeric(data$experimentalmasstocharge)
-      data$experimentalmasstocharge <- NULL
-   }
-   
-   if(!is.null(data$chargestate)){
-      data$chargeState <- data$chargestate
-      data$chargestate <- NULL
-   }
-   
-   if(!is.null(data$spectrumid)){
-      data$spectrumID <- data$spectrumid
-      data$spectrumid <- NULL
-   }
-   
-   if(!is.null(data$pre)){
-      data$Pre <- data$pre
-      data$pre <- NULL
-   }
-   
-   if(!is.null(data$post)){
[…]

### 2014-08-05 R/optimize_filter.r
speeding up optimization (3-fold) and changing tabs

@@ -5,103 +5,129 @@
 # It may make some (minor) sense to make it as MSnIDFilter-method though.
 
 
-.get_num_pep_for_fdr <- function(filterThresholds, msmsdata, filter, fdr.max, ...) 
+.get_num_pep_for_fdr <- function(filterThresholds, 
+                                    msmsdata, 
+                                    filter, 
+                                    fdr.max, 
+                                    ...) 
 {
-   filter <- update(filter, filterThresholds)
-   x <- evaluate_filter(msmsdata, filter, ...) # level should get here through ...
-   if(is.nan(x$fdr) || x$fdr > fdr.max){
-      return(rnorm(1,sd=0.001)) # 0 is bad because optimization does not move
-   }else{
-      return(x$n)
-   }
+    filter <- update(filter, filterThresholds)
+    # level should get here through ... (ellipsis)
+    x <- evaluate_filter(msmsdata, filter, ...)
+    if(is.nan(x$fdr) || x$fdr > fdr.max){
+        # 0 is bad because optimization does not move
+        return(rnorm(1,sd=0.001)) 
+    }else{
+        return(x$n)
+    }
 }
 
 
 .construct_optimization_grid <- function(filterObj, msnidObj, n.iter)
 {
-   #
-   # don't really like this -1 hack
-   n.iter.per.param <- round(n.iter^(1/length(filterObj))) - 1 
-   #
-   probs <- seq(0, 1, 1/n.iter.per.param)
-   eg <- expand.grid(
-      lapply(names(filterObj), 
-             function(arg) 
-                quantile(msnidObj[[arg]], probs, na.rm=T)))
-   colnames(eg) <- names(filterObj)
-   return(eg)
+    #
+    # todo: don't really like this -1 hack
+    # what is this for?
+    n.iter.per.param <- round(n.iter^(1/length(filterObj))) - 1 
+    #
+    probs <- seq(0, 1, 1/n.iter.per.param)
+    eg <- expand.grid(lapply(names(filterObj), 
+                                function(arg) 
+                                    quantile(msnidObj[[arg]], 
+                                                probs, na.rm=TRUE)))
+    colnames(eg) <- names(filterObj)
+    return(eg)
 }
 
 
 .optimize_filter.grid <- function(filterObj, msnidObj, fdr.max, level, n.iter)
 {
-   par.grid <- .construct_optimization_grid(filterObj, msnidObj, n.iter)
[…]

### 2014-08-05 R/MSnID-methods.R
speeding up id_quality. changing tabs from 3 to 4

@@ -3,106 +3,112 @@
 #----Peptide Sequence Handling--------------------------------------------------
 
 .get_clean_peptide_sequence <- function(peptide){
-   return(.strip_flanking_AAs(
-      .strip_modifications_from_peptide(peptide)))
+    return(
+        .strip_flanking_AAs(
+            .strip_modifications_from_peptide(peptide)))
 }
 
 .strip_modifications_from_peptide <- function(peptide)
 {
-   # INPUT: Peptide sequence with flanking AAs e.g. "K.APEP*TID{34.34}E%.-"
-   # OUTPUT: sequence with just AA symbols e.g. "K.APEPTIDE.-"
-   # clean from the mods
-   # make sure it removes "." in the middle of the sequence
-   aminoAcids <- "ARNDCEQGHILKMFPSTWYV"
-   extendedAAs <- "BUZXO" # http://en.wikipedia.org/wiki/Proteinogenic_amino_acid
-   # aminoAcids <- paste(LETTERS,collapse='') # hardcore solution
-   aminoAcids <- paste(aminoAcids, extendedAAs, sep='')
-   nonPeptideSequenceCharacter <- sprintf("[^%s.-]", aminoAcids)
-   peptide <- gsub( nonPeptideSequenceCharacter, '', peptide)
-   # stip "." in the middle of the sequence
-   # This wont touch the dots denoting cleavages, assuming (!) there is
-   # only one amino acid shown after the cleavage point.
-   peptide <- gsub( "(?<=.{2})\\.(?=.{2})", '', peptide, perl=TRUE)
-   return(peptide)
+    # INPUT: Peptide sequence with flanking AAs e.g. "K.APEP*TID{34.34}E%.-"
+    # OUTPUT: sequence with just AA symbols e.g. "K.APEPTIDE.-"
+    # clean from the mods
+    # make sure it removes "." in the middle of the sequence
+    aminoAcids <- "ARNDCEQGHILKMFPSTWYV"
+    # http://en.wikipedia.org/wiki/Proteinogenic_amino_acid
+    extendedAAs <- "BUZXO"
+    # aminoAcids <- paste(LETTERS,collapse='') # hardcore solution
+    aminoAcids <- paste(aminoAcids, extendedAAs, sep='')
+    nonPeptideSequenceCharacter <- sprintf("[^%s.-]", aminoAcids)
+    peptide <- gsub( nonPeptideSequenceCharacter, '', peptide)
+    # stip "." in the middle of the sequence
[…]

### 2014-08-05 demo/c_elegans.R
conforming the script with latest updates from mzID and MSnID object specificaitons

@@ -12,54 +12,27 @@ msnid <- read_mzIDs(msnid, mzids)
 #--- CHECKING WHAT IS INSIDE -----------
 head(psms(msnid))
 names(msnid)
-show(msnid) # Key columns are not set yet.
+show(msnid) 
 #---------------------------------------
-
-
-# --- UPDATES TO INCLUDE KEY COLUMNS ---
-# Note, the anticipated/suggested columns in the
-# peptide-to-spectrum matching results are:
-#    Peptide, Accession, isDecoy, 
-#    calculatedMassToCharge, experimentalMassToCharge, 
-#    spectrumFile, spectrumID
-msnid$Peptide <- paste(msnid$pre, msnid$pepseq, msnid$post, sep='.')
-msnid$Accession <- msnid$accession
-msnid$isDecoy <- msnid$isdecoy
-msnid$calculatedMassToCharge <- msnid$calculatedmasstocharge
-msnid$experimentalMassToCharge <- msnid$experimentalmasstocharge
-msnid$chargeState <- msnid$chargestate
-msnid$spectrumID <- msnid$spectrumid
-# tidying up. removing left-over columns
-msnid$accession <- NULL
-msnid$pepSeq <- NULL
-msnid$pre <- NULL
-msnid$post <- NULL
-msnid$idecoy <- NULL
-msnid$spectrumid <- NULL
-msnid$experimentalmasstocharge <- NULL
-msnid$calculatedmasstocharge <- NULL
-msnid$chargestate <- NULL
-show(msnid)
-#-------------------------------------
 
 
 # --- EXTRA INFO ABOUT PEPTIDE SEQUENCES ---
 msnid <- assess_termini(msnid, validCleavagePattern="[RK]\\.[^P]")
 msnid <- assess_missed_cleavages(msnid, missedCleavagePattern="[KR](?=[^P$])")
 # visualize distributions of missed cleavages as an example
-pepCleav <- unique(psms(msnid)[,c("NumMissCleavages", "isDecoy", "Peptide")])
-pepCleav <- as.data.frame(table(pepCleav[,c("NumMissCleavages", "isDecoy")]))
+pepCleav <- unique(psms(msnid)[,c("numMissCleavages", "isDecoy", "peptide")])
+pepCleav <- as.data.frame(table(pepCleav[,c("numMissCleavages", "isDecoy")]))
 library("ggplot2")
[…]

### 2014-08-05 R/MSnID-methods.R
changing a couple of variable names from CamelCase to camelCase

@@ -64,7 +64,7 @@ setMethod(
     definition=function(.Object, missedCleavagePattern)
     {
         .check_column(.Object, "peptide")
-        .Object@psms$NumMissCleavages <-
+        .Object@psms$numMissCleavages <-
             .assess_missed_cleavages(as.character(.Object@psms$peptide),
                                     missedCleavagePattern)
         return(.Object)
@@ -104,7 +104,7 @@ setMethod(
     signature("MSnID"),
     definition=function(.Object, validCleavagePattern)
     {
-        .Object@psms$NumIrregCleavages <- 
+        .Object@psms$numIrregCleavages <- 
             .assess_termini(as.character(.Object@psms$peptide), 
                             validCleavagePattern)
         return(.Object)

### 2014-08-06 R/MSnID-methods.R
.id_quality now returns a vector with names instead of a list

@@ -242,7 +242,7 @@ setMethod(
         return(list(fdr=NaN, n=0))
     # if there are normal matches, proceed
     fdr <- decoy/(n-decoy)
-    return(list(fdr=fdr, n=n))
+    return(c(fdr=fdr, n=n))
 }
 
 setMethod(
@@ -418,7 +418,8 @@ setMethod(
         for(i in c("PSM", "peptide", "accession")){
             try({
                 temp <- .id_quality(object, i)
-                cat("#", i, "s: ", temp$n, " at ", signif(100*temp$fdr, 2), 
+                cat("#", i, "s: ", temp['n'], " at ", 
+                    signif(100*temp['fdr'], 2), 
                     " % FDR", '\n', sep='')
                 }, 
                 silent=TRUE)

### 2014-08-06 R/optimize_filter.r

@@ -14,11 +14,11 @@
     filter <- update(filter, filterThresholds)
     # level should get here through ... (ellipsis)
     x <- evaluate_filter(msmsdata, filter, ...)
-    if(is.nan(x$fdr) || x$fdr > fdr.max){
+    if(is.nan(x['fdr']) || x['fdr'] > fdr.max){
         # 0 is bad because optimization does not move
         return(rnorm(1,sd=0.001)) 
     }else{
-        return(x$n)
+        return(x['n'])
     }
 }
 

### 2014-08-06 R/AllGenerics.R
deleted unused functions/methods

@@ -1,43 +1,24 @@
 
-setGeneric("set_filter",
-            function(.Object, ...) standardGeneric("set_filter"))
 
 setGeneric("assess_missed_cleavages",
             function(.Object, missedCleavagePattern="[KR](?=[^P$])") 
                 standardGeneric("assess_missed_cleavages"))
 
-# this may be an odd way to handle assess_termini
 setGeneric("assess_termini", 
             function(.Object, validCleavagePattern="[KR]\\.[^P]") 
                 standardGeneric("assess_termini"))
 
-setGeneric("get_accession_fdr", 
-            function(.Object, ...) standardGeneric("get_accession_fdr"))
+setGeneric("accessions", 
+            function(.Object) standardGeneric("accessions"))
 
-setGeneric("get_peptide_fdr", 
-            function(.Object, ...) standardGeneric("get_peptide_fdr"))
-
-setGeneric("get_psm_fdr", 
-            function(.Object, ...) standardGeneric("get_psm_fdr"))
-
-setGeneric("get_accessions", 
-            function(.Object) standardGeneric("get_accessions"))
-
-setGeneric("get_peptides", 
-            function(.Object) standardGeneric("get_peptides"))
-
-
-# setGeneric("read_mzIDs", 
-#            function(.Object, mzids) standardGeneric("read_mzIDs"))
+setGeneric("peptides", 
+            function(.Object) standardGeneric("peptides"))
 
 setGeneric("read_mzIDs", 
             function(object, mzids) standardGeneric("read_mzIDs"))
 
-
-
 setGeneric("apply_filter", 
             function(.Object, .Filter) standardGeneric("apply_filter"))
-
 
 setGeneric("evaluate_filter", 
             function(.Object, filter, level=c("PSM", "Peptide", "Accession"))

### 2014-08-06 R/MSnID-methods.R

@@ -118,58 +118,8 @@ setMethod(
 
 
 
-
-
-#----FDRs and unique peptides and accessions------------------------------------
 setMethod(
-    "get_psm_fdr", 
-    signature("MSnID"),
-    definition=function(.Object)
-    {
-        isDecoy <- table(factor(.Object@psms$isDecoy, 
-                                levels=c(FALSE, TRUE)))
-        fdr <- isDecoy["TRUE"]/isDecoy["FALSE"]
-        return(fdr)
-    }
-)
-
-
-
-
-setMethod(
-    "get_peptide_fdr", 
-    signature("MSnID"),
-    definition=function(.Object)
-    {
-        peptideDecoy <- unique(subset(.Object@psms, 
-                                        select=c("peptide","isDecoy")))
-        isDecoy <- table(factor(peptideDecoy$isDecoy, levels=c(FALSE, TRUE)))
-        fdr <- isDecoy["TRUE"]/isDecoy["FALSE"]
-        return(fdr)
-    }
-)
-
-
-
-setMethod(
-    "get_accession_fdr", 
-    signature("MSnID"),
-    definition=function(.Object)
-    {
-        accessionDecoy <- unique(subset(.Object@psms, 
-                                        select=c("accession","isDecoy")))
-        isDecoy <- table(factor(accessionDecoy$isDecoy, levels=c(FALSE, TRUE)))
-        fdr <- isDecoy["TRUE"]/isDecoy["FALSE"]
-        return(fdr)
-    }
-)
-
-
-
-
-# todo: change to peptides()
-setMethod(
-    "get_peptides", 
+    "peptides", 
     "MSnID",
     definition=function(.Object)
     {
@@ -179,16 +129,14 @@ setMethod(
 
 
 
-# todo: change to accessions()
 setMethod(
-    "get_accessions", 
+    "accessions", 
     "MSnID",
     definition=function(.Object)
     {
         unique(as.character(.Object@psms$accession))
     }
 )
-#-------------------------------------------------------------------------------
 
 
 

### 2014-08-06 R/MSnIDFilter-methods.R

@@ -1,20 +1,6 @@
 
 
 
-# setMethod("set_filter", 
-#             # dispatch just on the first argument
-#             signature(.Object="MSnIDFilter"), 
-#             definition=function(.Object, 
-#                                 parName,   # MUST be character
-#                                 comparison=NULL,
-#                                 threshold=NULL)
-#             {
-#                 parNames <- sapply(.Object@filterList, "[[", "name")
-#                 parNames <- sapply(.Object@filterList, "[[", "name")
-#             }
-# )
-
-
 .is_filterList_valid <- function(filterList)
 {
     entries=c("comparison","threshold")

### 2014-08-06 R/AllGenerics.R
protein functions. sibling to accessons

@@ -11,6 +11,9 @@ setGeneric("assess_termini",
 setGeneric("accessions", 
             function(.Object) standardGeneric("accessions"))
 
+setGeneric("proteins", 
+           function(.Object) standardGeneric("proteins"))
+
 setGeneric("peptides", 
             function(.Object) standardGeneric("peptides"))
 

### 2014-08-07 R/MSnID-methods.R
get_peptide and get_accessions to simply peptides and accessions (and proteins)

@@ -138,6 +138,14 @@ setMethod(
     }
 )
 
+setMethod(
+   "proteins", 
+   "MSnID",
+   definition=function(.Object)
+   {
+      accessions(.Object)
+   }
+)
 
 
 

### 2014-08-07 R/AllGenerics.R
switching .Object to object

@@ -1,48 +1,48 @@
 
 
 setGeneric("assess_missed_cleavages",
-            function(.Object, missedCleavagePattern="[KR](?=[^P$])") 
+            function(object, missedCleavagePattern="[KR](?=[^P$])") 
                 standardGeneric("assess_missed_cleavages"))
 
 setGeneric("assess_termini", 
-            function(.Object, validCleavagePattern="[KR]\\.[^P]") 
+            function(object, validCleavagePattern="[KR]\\.[^P]") 
                 standardGeneric("assess_termini"))
 
 setGeneric("accessions", 
-            function(.Object) standardGeneric("accessions"))
+            function(object) standardGeneric("accessions"))
 
 setGeneric("proteins", 
-           function(.Object) standardGeneric("proteins"))
+           function(object) standardGeneric("proteins"))
 
 setGeneric("peptides", 
-            function(.Object) standardGeneric("peptides"))
+            function(object) standardGeneric("peptides"))
 
 setGeneric("read_mzIDs", 
             function(object, mzids) standardGeneric("read_mzIDs"))
 
 setGeneric("apply_filter", 
-            function(.Object, .Filter) standardGeneric("apply_filter"))
+            function(object, .Filter) standardGeneric("apply_filter"))
 
 setGeneric("evaluate_filter", 
-            function(.Object, filter, level=c("PSM", "Peptide", "Accession"))
+            function(object, filter, level=c("PSM", "Peptide", "Accession"))
                 standardGeneric("evaluate_filter"))
 
 setGeneric("id_quality",
-            function(.Object, 
+            function(object, 
                         filter=NULL, 
                         level=c("PSM", "Peptide", "Accession")) 
                 standardGeneric("id_quality"))
 
 setGeneric("correct_peak_selection", 
-            function(.Object) 
+            function(object) 
                 standardGeneric("correct_peak_selection"))
 
 setGeneric("mass_measurement_error",
-            function(.Object) 
+            function(object) 
                 standardGeneric("mass_measurement_error"))
 
 setGeneric("recalibrate",
-            function(.Object) 
+            function(object) 
                 standardGeneric("recalibrate"))
 
 setGeneric("optimize_filter",
@@ -50,7 +50,7 @@ setGeneric("optimize_filter",
                 stadardGeneric("optimize_filter"))
 
 setGeneric("psms",
-            function(.Object) standardGeneric("psms"))
+            function(object) standardGeneric("psms"))
 
[…]

### 2014-08-07 R/MSnID-methods.R

@@ -61,13 +61,13 @@
 setMethod(
     "assess_missed_cleavages", 
     signature("MSnID"),
-    definition=function(.Object, missedCleavagePattern)
+    definition=function(object, missedCleavagePattern)
     {
-        .check_column(.Object, "peptide")
-        .Object@psms$numMissCleavages <-
-            .assess_missed_cleavages(as.character(.Object@psms$peptide),
+        .check_column(object, "peptide")
+        object@psms$numMissCleavages <-
+            .assess_missed_cleavages(as.character(object@psms$peptide),
                                     missedCleavagePattern)
-        return(.Object)
+        return(object)
     }
 )
 
@@ -102,12 +102,12 @@ setMethod(
 setMethod(
     "assess_termini", 
     signature("MSnID"),
-    definition=function(.Object, validCleavagePattern)
+    definition=function(object, validCleavagePattern)
     {
-        .Object@psms$numIrregCleavages <- 
-            .assess_termini(as.character(.Object@psms$peptide), 
+        object@psms$numIrregCleavages <- 
+            .assess_termini(as.character(object@psms$peptide), 
                             validCleavagePattern)
-        return(.Object)
+        return(object)
     }
 )
 
@@ -121,9 +121,9 @@ setMethod(
 setMethod(
     "peptides", 
     "MSnID",
-    definition=function(.Object)
+    definition=function(object)
     {
-        unique(as.character(.Object@psms$peptide))
+        unique(as.character(object@psms$peptide))
     }
 )
 
@@ -132,18 +132,18 @@ setMethod(
 setMethod(
     "accessions", 
     "MSnID",
-    definition=function(.Object)
+    definition=function(object)
     {
-        unique(as.character(.Object@psms$accession))
+        unique(as.character(object@psms$accession))
     }
 )
 
 setMethod(
    "proteins", 
    "MSnID",
-   definition=function(.Object)
+   definition=function(object)
    {
-      accessions(.Object)
+      accessions(object)
    }
 )
 
@@ -157,41 +157,41 @@ setMethod(
 #----Filter---------------------------------------------------------------------
 setMethod(
     "apply_filter", 
-    signature(.Object="MSnID", .Filter="character"),
-    definition=function(.Object, .Filter)
+    signature(object="MSnID", .Filter="character"),
+    definition=function(object, .Filter)
     {
         # the infamous eval(parse(text=...))
[…]

### 2014-08-07 R/MSnIDFilter-methods.R

@@ -10,11 +10,11 @@
 
 
 # internal
-.get_filterString <- function(.Object, precision=2)
+.get_filterString <- function(object, precision=2)
 {
     # 1) assuming names in the list match names of list elements
     # 2) assuming the names are in the rigth order
-    res <- lapply(.Object@filterList, 
+    res <- lapply(object@filterList, 
                     within, 
                     {threshold <- signif(threshold, 
                                         get("precision", parent.frame(n=4)))})
@@ -26,11 +26,11 @@
 }
 
 
-.get_filterValues <- function(.Object, precision=2)
+.get_filterValues <- function(object, precision=2)
 {
     # 1) assuming names in the list match names of list elements
     # 2) assuming the names are in the rigth order
-    res <- sapply(.Object@filterList, "[[", "threshold") 
+    res <- sapply(object@filterList, "[[", "threshold") 
     return(res)
 }
 

### 2014-08-07 R/AllGenerics.R
.Filter to filterObj

@@ -21,7 +21,7 @@ setGeneric("read_mzIDs",
             function(object, mzids) standardGeneric("read_mzIDs"))
 
 setGeneric("apply_filter", 
-            function(object, .Filter) standardGeneric("apply_filter"))
+            function(msnidObj, filterObj) standardGeneric("apply_filter"))
 
 setGeneric("evaluate_filter", 
             function(object, filter, level=c("PSM", "Peptide", "Accession"))

### 2014-08-07 R/MSnID-methods.R

@@ -157,24 +157,22 @@ setMethod(
 #----Filter---------------------------------------------------------------------
 setMethod(
     "apply_filter", 
-    signature(object="MSnID", .Filter="character"),
-    definition=function(object, .Filter)
+    signature(msnidObj="MSnID", filterObj="character"),
+    definition=function(msnidObj, filterObj)
     {
-        # the infamous eval(parse(text=...))
-        exprssn <- parse(text=.Filter, srcfile=NULL, keep.source=FALSE)
-        #object@psms <- subset(object@psms, eval(exprssn))
-        object@psms <- object@psms[eval(exprssn),]
-        return(object)
+        exprssn <- parse(text=filterObj, srcfile=NULL, keep.source=FALSE)
+        msnidObj@psms <- msnidObj@psms[eval(exprssn),]
+        return(msnidObj)
     }
 )
 
 setMethod(
     "apply_filter", 
-    signature(object="MSnID", .Filter="MSnIDFilter"),
-    definition=function(object, .Filter)
+    signature(msnidObj="MSnID", filterObj="MSnIDFilter"),
+    definition=function(msnidObj, filterObj)
     {
-        filterString <- as(.Filter, "character")
-        return(apply_filter(object, filterString))
+        filterString <- as(filterObj, "character")
+        return(apply_filter(msnidObj, filterString))
     }
 )
 

### 2014-08-07 R/AllGenerics.R
.Filter to filterObj and .Data to msnidObj

@@ -46,7 +46,7 @@ setGeneric("recalibrate",
                 standardGeneric("recalibrate"))
 
 setGeneric("optimize_filter",
-            function(.Filter, .Data, fdr.max, method, level, n.iter)
+            function(filterObj, msnidObj, fdr.max, method, level, n.iter)
                 stadardGeneric("optimize_filter"))
 
 setGeneric("psms",

### 2014-08-07 R/optimize_filter.r

@@ -76,54 +76,56 @@
 
 
 setMethod("optimize_filter",
-            signature(.Filter="MSnIDFilter", .Data="MSnID"),
-            definition=function(.Filter, .Data, fdr.max, method, level, n.iter)
+            signature(filterObj="MSnIDFilter", msnidObj="MSnID"),
+            definition=function(filterObj, msnidObj, fdr.max, 
+                                   method, level, n.iter)
             {
-                .optimize_filter.memoized(.Filter, .Data, 
+                .optimize_filter.memoized(filterObj, msnidObj, 
                                             fdr.max, method, level, n.iter)
             }
 )
 
 
 
-.optimize_filter <- function(.Filter, .Data, fdr.max, method, level, n.iter)
+.optimize_filter <- function(filterObj, msnidObj, 
+                               fdr.max, method, level, n.iter)
 {
     method <- match.arg(method, choices=c("Grid", "Nelder-Mead", "SANN"))
     level <- match.arg(level, choices=c("PSM", "peptide", "accession"))
     #
-    # subset .Data to only relevant columns
+    # subset msnidObj to only relevant columns
     if(level == "PSM"){
-        .Data@psms <- 
-            .Data@psms[,c("isDecoy", names(filtObj)), with=FALSE]
+       msnidObj@psms <- 
+           msnidObj@psms[,c("isDecoy", names(filtObj)), with=FALSE]
     }else{
-        .Data@psms <- 
-            .Data@psms[,c("isDecoy", level, names(filtObj)), with=FALSE]
+       msnidObj@psms <- 
+           msnidObj@psms[,c("isDecoy", level, names(filtObj)), with=FALSE]
     }
     # substitute Peptide and or Accession with integers
-    .Data@psms[[level]] <- as.integer(as.factor(.Data@psms[[level]]))
+    msnidObj@psms[[level]] <- as.integer(as.factor(msnidObj@psms[[level]]))
     #
     if(method == "Grid"){
         if(.Platform$OS.type == "unix"){
             optimFilter <- 
-                .optimize_filter.grid.mclapply(.Filter, .Data, 
+                .optimize_filter.grid.mclapply(filterObj, msnidObj, 
                                                 fdr.max, level, n.iter)
         }else{
             # yet to implement effective parallelization on Windows
             optimFilter <- 
-                .optimize_filter.grid(.Filter, .Data, 
+                .optimize_filter.grid(filterObj, msnidObj, 
[…]

### 2014-08-07 R/optimize_filter.r
cleaning comments

@@ -1,9 +1,4 @@
-# Here will be the methods for handling filter optimization.
-# Since they apply both to MSnID and MSnIDFilter objects 
-# it probably make sense not to make these methods exclusively
-# for either of the classes. 
-# It may make some (minor) sense to make it as MSnIDFilter-method though.
-
+# methods and functions for handling filter optimization.
 
 .get_num_pep_for_fdr <- function(filterThresholds, 
                                     msmsdata, 
@@ -25,11 +20,7 @@
 
 .construct_optimization_grid <- function(filterObj, msnidObj, n.iter)
 {
-    #
-    # todo: don't really like this -1 hack
-    # what is this for?
     n.iter.per.param <- round(n.iter^(1/length(filterObj))) - 1 
-    #
     probs <- seq(0, 1, 1/n.iter.per.param)
     eg <- expand.grid(lapply(names(filterObj), 
                                 function(arg) 

### 2014-08-07 R/AllGenerics.R
changing case in Peptides and Accessions to low

@@ -24,13 +24,13 @@ setGeneric("apply_filter",
             function(msnidObj, filterObj) standardGeneric("apply_filter"))
 
 setGeneric("evaluate_filter", 
-            function(object, filter, level=c("PSM", "Peptide", "Accession"))
+            function(object, filter, level=c("PSM", "peptide", "accession"))
                 standardGeneric("evaluate_filter"))
 
 setGeneric("id_quality",
             function(object, 
                         filter=NULL, 
-                        level=c("PSM", "Peptide", "Accession")) 
+                        level=c("PSM", "peptide", "accession")) 
                 standardGeneric("id_quality"))
 
 setGeneric("correct_peak_selection", 

### 2014-08-07 R/optimize_filter.r
fixing named vector accessing issue

@@ -9,11 +9,11 @@
     filter <- update(filter, filterThresholds)
     # level should get here through ... (ellipsis)
     x <- evaluate_filter(msmsdata, filter, ...)
-    if(is.nan(x['fdr']) || x['fdr'] > fdr.max){
+    if(is.nan(x[['fdr']]) || x[['fdr']] > fdr.max){
         # 0 is bad because optimization does not move
         return(rnorm(1,sd=0.001)) 
     }else{
-        return(x['n'])
+        return(x[['n']])
     }
 }
 
@@ -87,10 +87,10 @@ setMethod("optimize_filter",
     # subset msnidObj to only relevant columns
     if(level == "PSM"){
        msnidObj@psms <- 
-           msnidObj@psms[,c("isDecoy", names(filtObj)), with=FALSE]
+           msnidObj@psms[,c("isDecoy", names(filterObj)), with=FALSE]
     }else{
        msnidObj@psms <- 
-           msnidObj@psms[,c("isDecoy", level, names(filtObj)), with=FALSE]
+           msnidObj@psms[,c("isDecoy", level, names(filterObj)), with=FALSE]
     }
     # substitute Peptide and or Accession with integers
     msnidObj@psms[[level]] <- as.integer(as.factor(msnidObj@psms[[level]]))

### 2014-08-07 inst/unitTests/test_apply_filter.R
switch from list to named vector

@@ -5,11 +5,11 @@ test_apply_filter_1 <- function(){
     # checking string filter
     out <- apply_filter(msnidObj, "`MS-GF:SpecEValue` < 10^-10")
     checkEquals(id_quality(out, level="PSM"),
-                list(fdr=0.000285103349964362, n=7017))
+                c(fdr=0.000285103349964362, n=7017))
     checkEquals(id_quality(out, level="peptide"),
-                list(fdr=0.000778816199376947, n=2570))
+                c(fdr=0.000778816199376947, n=2570))
     checkEquals(id_quality(out, level="accession"),
-                list(fdr=0.00193986420950533, n=1033))
+                c(fdr=0.00193986420950533, n=1033))
 }
 
 
@@ -27,11 +27,11 @@ test_apply_filter_2 <- function(){
     out <- apply_filter(msnidObj, filtObj)
 
     checkEquals(id_quality(out, level="PSM"),
-                list(fdr=0.0011079650375477, n=8132))
+                c(fdr=0.0011079650375477, n=8132))
     checkEquals(id_quality(out, level="peptide"),
-                list(fdr=0.0013550135501355, n=2956))
+                c(fdr=0.0013550135501355, n=2956))
     checkEquals(id_quality(out, level="accession"),
-                list(fdr=0.00801424755120214, n=1132))
+                c(fdr=0.00801424755120214, n=1132))
 }
 
 

### 2014-08-07 R/AllGenerics.R
fixing spaces

@@ -12,7 +12,7 @@ setGeneric("accessions",
             function(object) standardGeneric("accessions"))
 
 setGeneric("proteins", 
-           function(object) standardGeneric("proteins"))
+            function(object) standardGeneric("proteins"))
 
 setGeneric("peptides", 
             function(object) standardGeneric("peptides"))

### 2014-08-07 R/MSnID-methods.R

@@ -139,12 +139,12 @@ setMethod(
 )
 
 setMethod(
-   "proteins", 
-   "MSnID",
-   definition=function(object)
-   {
-      accessions(object)
-   }
+    "proteins", 
+    "MSnID",
+    definition=function(object)
+    {
+        accessions(object)
+    }
 )
 
 

### 2014-08-07 R/MSnID-methods.R
fix empty message

@@ -297,7 +297,7 @@ setReplaceMethod(
     definition=function(object, value)
     {
         misCol <- setdiff(.mustBeColumns, colnames(value))
-        if(!is.null(misCol) & interactive()){
+        if((length(misCol) > 0) & interactive()){
             promptStr <- 
                 paste("The data.frame is missing the following columns:\n",
                     paste(strwrap(paste(misCol, collapse=', ')), collapse='\n'),

### 2014-08-07 R/MSnID-methods.R
cleaning comments

@@ -401,13 +401,6 @@ setMethod(
 
 
 #-------------------------------------------------------------------------------
-# Getters and setters modelled after
-# library(Biobase)
-# getMethod("$", "eSet") # tricky
-# getMethod("$<-", "eSet") # straight
-# getMethod("[[", "eSet")
-# getMethod("[[<-", "eSet")
-
 # Note, get method in Biobase is a bit different
 # library(Biobase)
 # getMethod("$", "eSet")
@@ -448,17 +441,6 @@ setMethod("$<-", "MSnID",
         return(x)
     }
 )
-
-
-# setMethod("$<-", "MSnID",
-#     definition=function(x, name, value)
-#     {
-#         x@psms <- eval(substitute(x@psms[,NAME_ARG:=value],
-#                                   list(NAME_ARG=name)))
-#         return(x)
-#     }
-# )
-
 
 
 setMethod("[[<-", "MSnID",

### 2014-08-08 R/MSnID-methods.R
finalizing $ and $<- operators

@@ -418,30 +418,13 @@ setMethod("[[", "MSnID",
             }
 )
 
-
-
-# todo: double check set $<- method
-# todo: the current problem is that := assignment by reference
-# still does not work properly (copies by value). Therefore
-# will use either set or find smarter way to use :=
-
-# # data.frame style
-# setMethod("$<-", "MSnID",
-#           definition=function(x, name, value)
-#           {
-#              x@psms[[name]] <- value
-#              return(x)
-#           }
-# )
 setMethod("$<-", "MSnID",
     definition=function(x, name, value)
     {
-        x@psms <- eval(substitute(x@psms[,NAME_ARG:=value],
-                                    list(NAME_ARG=name)))
+        x@psms[[name]] <- value
         return(x)
     }
 )
-
 
 setMethod("[[<-", "MSnID",
             definition=function(x, i, j, ..., value)

### 2014-08-08 R/MSnID-methods.R
cleaning comments

@@ -117,7 +117,7 @@ setMethod(
 
 
 
-
+#--- Accessors -----------------------------------------------------------------
 setMethod(
     "peptides", 
     "MSnID",
@@ -146,7 +146,7 @@ setMethod(
         accessions(object)
     }
 )
-
+#-------------------------------------------------------------------------------
 
 
 
@@ -176,9 +176,6 @@ setMethod(
     }
 )
 
-
-
-# todo: change to returning a vector with names instead of a list
 .id_quality <- function(object, .Level=c("PSM", "peptide", "accession"))
 {
     #
@@ -401,9 +398,6 @@ setMethod(
 
 
 #-------------------------------------------------------------------------------
-# Note, get method in Biobase is a bit different
-# library(Biobase)
-# getMethod("$", "eSet")
 setMethod("$", "MSnID",
             definition=function(x, name)
             {

### 2014-08-08 R/AllClasses.R
formatting

@@ -9,9 +9,6 @@ setClass(Class="MSnID",
 )
 
 
-
-
-# partitions and parameter transforms to be added to filter class
 setClass(Class="MSnIDFilter",
             representation(
                 filterList="list",

### 2014-08-08 R/optimize_filter.r
fixing legacy of switching from list to named vector

@@ -116,7 +116,7 @@ setMethod("optimize_filter",
                             level = level,
                             method = method,
                             control=list(fnscale=-1, maxit=n.iter))
-        optimFilter <- update(filterObj, optim.out$par)
+        optimFilter <- update(filterObj, as.numeric(optim.out$par))
     }
     return(optimFilter)
 }

### 2014-09-20 R/MSnID-methods.R
changing cat to message

@@ -32,7 +32,7 @@
 {
     # INPUT: Peptide sequence with flanking AAs e.g. "K.APEP*TID{34.34}E%.-"
     # OUPUT: Peptide without flanking "APEP*TID{34.34}E%"
-    # 
+    #
     # this is quicker, but may not be safe if there are some mods with dots
     # stopifnot(nchar(peptide) - nchar(gsub("\\.","",peptide)) == 2)
     # this one is safer, but likely to be slower
@@ -51,15 +51,15 @@
     # the fastest approach
     peptide <- .get_clean_peptide_sequence(peptide)
     # now no flanking AAs, no mods
-    numberOfMissedCleavages <- 
-        nchar(peptide) - 
+    numberOfMissedCleavages <-
+        nchar(peptide) -
         nchar(gsub(missedCleavagePattern, "", peptide, perl=TRUE))
     return( numberOfMissedCleavages )
 }
 
 
 setMethod(
-    "assess_missed_cleavages", 
+    "assess_missed_cleavages",
     signature("MSnID"),
     definition=function(object, missedCleavagePattern)
     {
@@ -73,12 +73,12 @@ setMethod(
 
 #---
 
-.assess_termini <- function(peptide, 
+.assess_termini <- function(peptide,
                             validCleavagePattern)
 {
     # "[RK]\\.[^P]" | "-\\."
     # It has to be sequence with flanking AAs e.g. K.XXXXXXXR.X
-    # first peptide needs to be cleaned on anything other then 
+    # first peptide needs to be cleaned on anything other then
     # 20 amino acids and . and -
     #
     # make sure all have flanked AAs (this line of code is redundant)
@@ -91,7 +91,7 @@ setMethod(
     is.C.valid <- grepl(sprintf("%s$", validCleavagePattern), peptide)
     is.protein.N.terminus <- grepl("^-\\.",peptide)
     is.protein.C.terminus <- grepl("\\.-$",peptide)
-    number.of.irregular.termini <- 
+    number.of.irregular.termini <-
         as.numeric(!(is.N.valid | is.protein.N.terminus)) + # valid N-term
         as.numeric(!(is.C.valid | is.protein.C.terminus))   # valid C-term
     return(number.of.irregular.termini)
@@ -100,12 +100,12 @@ setMethod(
 
 
 setMethod(
-    "assess_termini", 
+    "assess_termini",
     signature("MSnID"),
     definition=function(object, validCleavagePattern)
     {
[…]

### 2014-09-20 R/read_mzIDs.R

@@ -5,29 +5,29 @@
     key <- list(mzids)
     data <- loadCache(key)
     if (!is.null(data)){
-        cat("Loaded cached data\n")
-    }else{   
-        cat("Reading from mzIdentMLs ...\n")
+        message("Loaded cached data\n")
+    }else{
+        message("Reading from mzIdentMLs ...\n")
         data <- data.table(flatten(mzID(mzids), safeNames=FALSE))
         saveCache(data, key=key)
     }
 
     # Adding peptide identification in a conventional format:
-    # X.XXXXX.X, where pre and post amino acids separated by dots 
+    # X.XXXXX.X, where pre and post amino acids separated by dots
     # from the main sequence
     # This may be substituted in the future with more flexible
     # peptide S4 object
     data$peptide <- paste(data$pre, data$pepSeq, data$post, sep='.')
 
     # Columns that must be present:
-    # peptide, accession, isDecoy, calculatedMassToCharge, 
+    # peptide, accession, isDecoy, calculatedMassToCharge,
     # experimentalMassToCharge, chargeState, spectrumFile, spectrumID
     # There is no checking for these columns at this point.
     # If the mzIdentML file conformed with PSI standard, most likely
     # those columns are present.
     # There are checks for columns in some of the other functions
     # that rely on presence of particular columns.
-    
+
     return(data)
 }
 

### 2014-09-27 R/AllGenerics.R
Adds EnrichmentBrowser/ GOexpress/ MGFM/ MSnID/ rRDP/ simulatorZ/ to the repos.
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnID@94601 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -1,0 +1,56 @@
+
+
+setGeneric("assess_missed_cleavages",
+            function(object, missedCleavagePattern="[KR](?=[^P$])") 
+                standardGeneric("assess_missed_cleavages"))
+
+setGeneric("assess_termini", 
+            function(object, validCleavagePattern="[KR]\\.[^P]") 
+                standardGeneric("assess_termini"))
+
+setGeneric("accessions", 
+            function(object) standardGeneric("accessions"))
+
+setGeneric("proteins", 
+            function(object) standardGeneric("proteins"))
+
+setGeneric("peptides", 
+            function(object) standardGeneric("peptides"))
+
+setGeneric("read_mzIDs", 
+            function(object, mzids) standardGeneric("read_mzIDs"))
+
+setGeneric("apply_filter", 
+            function(msnidObj, filterObj) standardGeneric("apply_filter"))
+
+setGeneric("evaluate_filter", 
+            function(object, filter, level=c("PSM", "peptide", "accession"))
+                standardGeneric("evaluate_filter"))
+
+setGeneric("id_quality",
+            function(object, 
+                        filter=NULL, 
+                        level=c("PSM", "peptide", "accession")) 
+                standardGeneric("id_quality"))
+
+setGeneric("correct_peak_selection", 
+            function(object) 
+                standardGeneric("correct_peak_selection"))
+
+setGeneric("mass_measurement_error",
+            function(object) 
+                standardGeneric("mass_measurement_error"))
+
+setGeneric("recalibrate",
+            function(object) 
+                standardGeneric("recalibrate"))
+
+setGeneric("optimize_filter",
+            function(filterObj, msnidObj, fdr.max, method, level, n.iter)
+                stadardGeneric("optimize_filter"))
+
+setGeneric("psms",
+            function(object) standardGeneric("psms"))
+
+setGeneric("psms<-",
+            function(object, value) standardGeneric("psms<-"))

### 2014-09-27 R/read_mzIDs.R

@@ -1,0 +1,36 @@
+
+.read_mzIDs.memoized <- function(mzids)
+{
+    # Try to load cached data, if exists
+    key <- list(mzids)
+    data <- loadCache(key)
+    if (!is.null(data)){
+        message("Loaded cached data\n")
+    }else{
+        message("Reading from mzIdentMLs ...\n")
+        data <- data.table(flatten(mzID(mzids), safeNames=FALSE))
+        saveCache(data, key=key)
+    }
+
+    # Adding peptide identification in a conventional format:
+    # X.XXXXX.X, where pre and post amino acids separated by dots
+    # from the main sequence
+    # This may be substituted in the future with more flexible
+    # peptide S4 object
+    data$peptide <- paste(data$pre, data$pepSeq, data$post, sep='.')
+
+    # Columns that must be present:
+    # peptide, accession, isDecoy, calculatedMassToCharge,
+    # experimentalMassToCharge, chargeState, spectrumFile, spectrumID
+    # There is no checking for these columns at this point.
+    # If the mzIdentML file conformed with PSI standard, most likely
+    # those columns are present.
+    # There are checks for columns in some of the other functions
+    # that rely on presence of particular columns.
+
+    return(data)
+}
+
+
+
+

### 2014-09-27 demo/c_elegans.R

@@ -1,0 +1,166 @@
+
+
+#--- READING MZID FILE(S) ----------------
+library("MSnID")
+# start the project by providing work directory
+msnid <- MSnID(".")
+mzids <- system.file("extdata","c_elegans.mzid.gz",package="MSnID")
+msnid <- read_mzIDs(msnid, mzids)
+#---------------------------------------
+
+
+#--- CHECKING WHAT IS INSIDE -----------
+head(psms(msnid))
+names(msnid)
+show(msnid) 
+#---------------------------------------
+
+
+# --- EXTRA INFO ABOUT PEPTIDE SEQUENCES ---
+msnid <- assess_termini(msnid, validCleavagePattern="[RK]\\.[^P]")
+msnid <- assess_missed_cleavages(msnid, missedCleavagePattern="[KR](?=[^P$])")
+# visualize distributions of missed cleavages as an example
+pepCleav <- unique(psms(msnid)[,c("numMissCleavages", "isDecoy", "peptide")])
+pepCleav <- as.data.frame(table(pepCleav[,c("numMissCleavages", "isDecoy")]))
+library("ggplot2")
+ggplot(pepCleav, aes(x=numMissCleavages, y=Freq, fill=isDecoy)) + 
+    geom_bar(stat="identity", position="dodge") + 
+    ggtitle("Number of Missed Cleavages")
+# number of cysteins per peptide sequence as an example of sequence analysis
+msnid$NumCys <- sapply(lapply(strsplit(msnid$peptide,''),'==','C'),sum)
+# calculating peptide lengths
+msnid$pepLength <- nchar(msnid$peptide) - 4
+pepLen <- unique(psms(msnid)[,c("pepLength", "isDecoy", "peptide")])
+# distribution of peptide lengths in the dataset
+ggplot(pepLen, aes(x=pepLength, fill=isDecoy)) + 
+    geom_histogram(position='dodge', binwidth=3) +
+    ggtitle("Distribution on of Peptide Lengths")
+#------------------------------------------
+
+
+# --- TRIM THE DATA -----------------------
+# Let's take a look how filtering on irregular and
+# missed cleavages will affect the FDR.
+show(msnid)
+# 1. Leave only fully tryptic
+msnid <- apply_filter(msnid, "numIrregCleavages == 0")
+show(msnid)
+# 2. Retain peptides with at most 2 missed cleavages
+msnid <- apply_filter(msnid, "numMissCleavages <= 2")
+show(msnid)
+#-----------------------------------------
+
+
[…]

### 2014-09-27 inst/unitTests/test_optimize_filter.R

@@ -1,0 +1,48 @@
+library("MSnID")
+data(c_elegans)
+
+msnidObj$msmsScore <- -log10(msnidObj$`MS-GF:SpecEValue`)
+msnidObj$absParentMassErrorPPM <- abs(mass_measurement_error(msnidObj))
+
+filtObj <- MSnIDFilter(msnidObj)
+filtObj$absParentMassErrorPPM <- list(comparison="<", threshold=10.0)
+filtObj$msmsScore <- list(comparison=">", threshold=10.0)
+
+set.seed(0)
+filtObj.grid <- optimize_filter(filtObj, msnidObj, fdr.max=0.001, 
+                                method="Grid", level="peptide", 
+                                n.iter=500)
+
+set.seed(0)
+filtObj.nm <- optimize_filter(filtObj.grid, msnidObj, fdr.max=0.001, 
+                              method="Nelder-Mead", level="peptide", 
+                              n.iter=500)
+
+set.seed(0)
+filtObj.sann <- optimize_filter(filtObj.grid, msnidObj, fdr.max=0.001, 
+                              method="SANN", level="peptide", 
+                              n.iter=500)
+
+
+test_grid_optimization <- function(){
+    checkEquals(filtObj.grid@filterList$absParentMassErrorPPM$threshold,
+                3.99291882215223)
+    checkEquals(filtObj.grid@filterList$msmsScore$threshold,
+                9.04628489941093)
+}
+
+
+test_Nelder_Mead_optimization <- function(){
+    checkEquals(filtObj.nm@filterList$absParentMassErrorPPM$threshold,
+                3.81032565709881)
+    checkEquals(filtObj.nm@filterList$msmsScore$threshold,
+                8.87912408683179)
+}
+
+
+test_SANN_optimization <- function(){
+    checkEquals(filtObj.sann@filterList$absParentMassErrorPPM$threshold,
+                3.99800146330237)
+    checkEquals(filtObj.sann@filterList$msmsScore$threshold,
+                8.8787177674537)
+}

### 2014-09-27 R/AllClasses.R
Adds EnrichmentBrowser/ GOexpress/ MGFM/ MSnID/ rRDP/ simulatorZ/ to the repos.
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnID@94601 bc3139a8-67e5-0310-9ffc-ced21a209358

@@ -1,0 +1,17 @@
+
+
+.PROTON_MASS <- 1.007276466812
+
+setClass(Class="MSnID",
+            representation(
+                workDir="character", # working directory
+                psms="data.table") # peptide-to-spectra matches
+)
+
+
+setClass(Class="MSnIDFilter",
+            representation(
+                filterList="list",
+                validParNames="character")
+)
+

### 2014-09-27 R/MSnID-methods.R

@@ -1,0 +1,529 @@
+
+
+#----Peptide Sequence Handling--------------------------------------------------
+
+.get_clean_peptide_sequence <- function(peptide){
+    return(
+        .strip_flanking_AAs(
+            .strip_modifications_from_peptide(peptide)))
+}
+
+.strip_modifications_from_peptide <- function(peptide)
+{
+    # INPUT: Peptide sequence with flanking AAs e.g. "K.APEP*TID{34.34}E%.-"
+    # OUTPUT: sequence with just AA symbols e.g. "K.APEPTIDE.-"
+    # clean from the mods
+    # make sure it removes "." in the middle of the sequence
+    aminoAcids <- "ARNDCEQGHILKMFPSTWYV"
+    # http://en.wikipedia.org/wiki/Proteinogenic_amino_acid
+    extendedAAs <- "BUZXO"
+    # aminoAcids <- paste(LETTERS,collapse='') # hardcore solution
+    aminoAcids <- paste(aminoAcids, extendedAAs, sep='')
+    nonPeptideSequenceCharacter <- sprintf("[^%s.-]", aminoAcids)
+    peptide <- gsub( nonPeptideSequenceCharacter, '', peptide)
+    # stip "." in the middle of the sequence
+    # This wont touch the dots denoting cleavages, assuming (!) there is
+    # only one amino acid shown after the cleavage point.
+    peptide <- gsub( "(?<=.{2})\\.(?=.{2})", '', peptide, perl=TRUE)
+    return(peptide)
+}
+
+.strip_flanking_AAs <- function(peptide)
+{
+    # INPUT: Peptide sequence with flanking AAs e.g. "K.APEP*TID{34.34}E%.-"
+    # OUPUT: Peptide without flanking "APEP*TID{34.34}E%"
+    #
+    # this is quicker, but may not be safe if there are some mods with dots
+    # stopifnot(nchar(peptide) - nchar(gsub("\\.","",peptide)) == 2)
+    # this one is safer, but likely to be slower
+    stopifnot(all(grepl("^.\\..+\\..$", peptide)))
+    #
+    # remove flanking AAs
+    peptide <- substring( peptide, 3, nchar(peptide)-2)
+    #
+    return(peptide)
+}
+
+
+.assess_missed_cleavages <- function(peptide,
+                                        missedCleavagePattern)
+{
+    # the fastest approach
+    peptide <- .get_clean_peptide_sequence(peptide)
[…]

### 2014-09-27 R/MSnIDFilter-methods.R

@@ -1,0 +1,144 @@
+
+
+
+.is_filterList_valid <- function(filterList)
+{
+    entries=c("comparison","threshold")
+    all(sapply(filterList, function(v) all(entries %in% names(v))))
+}
+
+
+
+# internal
+.get_filterString <- function(object, precision=2)
+{
+    # 1) assuming names in the list match names of list elements
+    # 2) assuming the names are in the rigth order
+    res <- lapply(object@filterList, 
+                    within, 
+                    {threshold <- signif(threshold, 
+                                        get("precision", parent.frame(n=4)))})
+    res <- lapply(names(res), 
+                    function(v) 
+                        paste(c(list(v), res[[v]]), collapse=' '))
+    res <- paste(sprintf("(%s)", res), collapse=" & ")
+    return(res)
+}
+
+
+.get_filterValues <- function(object, precision=2)
+{
+    # 1) assuming names in the list match names of list elements
+    # 2) assuming the names are in the rigth order
+    res <- sapply(object@filterList, "[[", "threshold") 
+    return(res)
+}
+
+
+setAs("MSnIDFilter", "character",
+        def=function(from) .get_filterString(from, precision=Inf))
+
+
+setAs("MSnIDFilter", "numeric",
+        def=function(from) .get_filterValues(from, precision=Inf))
+
+setMethod("as.numeric", "MSnIDFilter",
+            definition=function(x, ...) as(x,"numeric"))
+
+setMethod("length", "MSnIDFilter",
+            definition=function(x) length(x@filterList))
+
+setMethod("names", "MSnIDFilter",
+            definition=function(x) names(x@filterList))
+
+setMethod("update", "MSnIDFilter",
+            definition=function(object, newThresholds)
+            {
+                stopifnot(length(object) == length(newThresholds))
+                for(i in seq_along(object))
+                    object@filterList[[i]]$threshold <- newThresholds[i]
+                return(object)
+            }
+)
+
+setMethod("show", "MSnIDFilter",
+            definition=function(object)
+            {
+                cat(class(object)," object\n",sep='')
+                cat(.get_filterString(object),'\n')
+            }
+)
+
+
+MSnIDFilter <- function(MSnIDObj, filterList=list())
+{
+    # validations supposed to be *before* the return
+    # let's allow only complete filters now
+    if(!is.list(filterList)){
+        stop(deparse(substitute(filterList)), " is not a list!")
+    }
+    if(!.is_filterList_valid(filterList)){
[…]

### 2014-09-27 R/optimize_filter.r

@@ -1,0 +1,126 @@
+# methods and functions for handling filter optimization.
+
+.get_num_pep_for_fdr <- function(filterThresholds, 
+                                    msmsdata, 
+                                    filter, 
+                                    fdr.max, 
+                                    ...) 
+{
+    filter <- update(filter, filterThresholds)
+    # level should get here through ... (ellipsis)
+    x <- evaluate_filter(msmsdata, filter, ...)
+    if(is.nan(x[['fdr']]) || x[['fdr']] > fdr.max){
+        # 0 is bad because optimization does not move
+        return(rnorm(1,sd=0.001)) 
+    }else{
+        return(x[['n']])
+    }
+}
+
+
+.construct_optimization_grid <- function(filterObj, msnidObj, n.iter)
+{
+    n.iter.per.param <- round(n.iter^(1/length(filterObj))) - 1 
+    probs <- seq(0, 1, 1/n.iter.per.param)
+    eg <- expand.grid(lapply(names(filterObj), 
+                                function(arg) 
+                                    quantile(msnidObj[[arg]], 
+                                                probs, na.rm=TRUE)))
+    colnames(eg) <- names(filterObj)
+    return(eg)
+}
+
+
+.optimize_filter.grid <- function(filterObj, msnidObj, fdr.max, level, n.iter)
+{
+    par.grid <- .construct_optimization_grid(filterObj, msnidObj, n.iter)
+    evaluations <- apply(X=par.grid, 
+                            MARGIN=1, 
+                            FUN=.get_num_pep_for_fdr, 
+                                msnidObj, 
+                                filterObj, 
+                                fdr.max, 
+                                level)
+    optim.pars <- par.grid[which.max(evaluations),]
+    newFilter <- update(filterObj, as.numeric(optim.pars))
+    return(newFilter)
+}
+
+
+.optimize_filter.grid.mclapply <- function(filterObj, msnidObj, 
+                                            fdr.max, level, n.iter)
+{
+    par.grid <- .construct_optimization_grid(filterObj, msnidObj, n.iter)
+    evaluations <- mclapply(seq_len(nrow(par.grid)), 
+                            function(i){
+                                .get_num_pep_for_fdr(par.grid[i,], 
+                                                        msnidObj, 
+                                                        filterObj, 
+                                                        fdr.max, 
+                                                        level)}, 
+                            mc.cores=detectCores())
[…]

### 2014-09-27 inst/unitTests/test_apply_filter.R

@@ -1,0 +1,37 @@
+library("MSnID")
+data(c_elegans)
+
+test_apply_filter_1 <- function(){
+    # checking string filter
+    out <- apply_filter(msnidObj, "`MS-GF:SpecEValue` < 10^-10")
+    checkEquals(id_quality(out, level="PSM"),
+                c(fdr=0.000285103349964362, n=7017))
+    checkEquals(id_quality(out, level="peptide"),
+                c(fdr=0.000778816199376947, n=2570))
+    checkEquals(id_quality(out, level="accession"),
+                c(fdr=0.00193986420950533, n=1033))
+}
+
+
+test_apply_filter_2 <- function(){
+    # checking object-based filter
+
+    msnidObj$msmsScore <- -log10(msnidObj$`MS-GF:SpecEValue`)
+    msnidObj$mzError <- abs(msnidObj$experimentalMassToCharge - 
+                            msnidObj$calculatedMassToCharge)
+    # setting up filter object
+    filtObj <- MSnIDFilter(msnidObj)
+    filtObj$msmsScore <- list(comparison=">", threshold=9.0)
+    filtObj$mzError <- list(comparison="<", threshold=0.2)
+
+    out <- apply_filter(msnidObj, filtObj)
+
+    checkEquals(id_quality(out, level="PSM"),
+                c(fdr=0.0011079650375477, n=8132))
+    checkEquals(id_quality(out, level="peptide"),
+                c(fdr=0.0013550135501355, n=2956))
+    checkEquals(id_quality(out, level="accession"),
+                c(fdr=0.00801424755120214, n=1132))
+}
+
+

### 2014-09-27 inst/unitTests/test_data_load.R

@@ -1,0 +1,16 @@
+library("MSnID")
+msnid2 <- MSnID(".")
+mzids <- system.file("extdata","c_elegans.mzid.gz",package="MSnID")
+msnid2 <- read_mzIDs(msnid2, mzids)
+unlink(".Rcache", recursive=TRUE)
+
+
+# Note, the psms slot is data.table
+# the .internal.selfref is session specific.
+# So I'll compare psms as data.frame only.
+
+test_data_load <- function() {
+    # now, check if it is what it supposed to be
+    data(c_elegans)
+    checkIdentical(psms(msnidObj), psms(msnid2))
+}

### 2014-09-27 inst/unitTests/test_irreg_cleavages.R

@@ -1,0 +1,19 @@
+library("MSnID")
+library("digest")
+data(c_elegans)
+msnidObj <- assess_termini(msnidObj, validCleavagePattern="[KR]\\.[^P]")
+
+test_check_name_presence <- function(){
+    checkTrue("numIrregCleavages" %in% names(msnidObj))
+}
+
+test_check_type <- function(){
+    checkTrue(is.numeric(msnidObj$numIrregCleavages))
+}
+
+test_check_result <- function(){
+    checkIdentical(digest(msnidObj$numIrregCleavages),
+                   "e0c4ab882ffb0c626c53ab801dc5a7e3")
+}
+
+

### 2014-09-27 inst/unitTests/test_missed_cleavages.R

@@ -1,0 +1,20 @@
+library("MSnID")
+library("digest")
+data(c_elegans)
+msnidObj <- assess_missed_cleavages(msnidObj, 
+                                    missedCleavagePattern="[KR](?=[^P$])")
+
+
+test_check_name_presence <- function(){
+    checkTrue("numMissCleavages" %in% names(msnidObj))
+}
+
+test_check_type <- function(){
+    checkTrue(is.numeric(msnidObj$numMissCleavages))
+}
+
+test_check_result <- function(){
+    checkIdentical(digest(msnidObj$numMissCleavages),
+                   "5d6f1fbd250a5ee10b0ad9d9c171ed90")
+}
+

### 2014-09-27 tests/runTests.R

@@ -1,0 +1,1 @@
+BiocGenerics:::testPackage("MSnID")

### 2017-10-05 R/optimize_filter.r
switch from mclapply to parLapply

@@ -88,6 +88,36 @@
 }
 
 
+.optimize_filter.grid.parLapply <- function(filterObj, msnidObj,
+                                           fdr.max, level, n.iter, mc.cores)
+{
+    par.grid <- .construct_optimization_grid(filterObj, msnidObj, n.iter)
+    cl <- makeCluster(mc.cores)
+    # clusterExport(cl, 
+    #                         list(".get_num_pep_for_fdr"),
+    #                         envir = "package:MSnID")
+    clusterEvalQ(cl, library("MSnID"))
+    evaluations <- parLapply(cl,
+                             seq_len(nrow(par.grid)),
+                             function(i){
+                                .get_num_pep_for_fdr(par.grid[i,],
+                                                     msnidObj,
+                                                     filterObj,
+                                                     fdr.max,
+                                                     level)})
+    stopCluster(cl)
+    evaluations <- round(unlist(evaluations))
+    if(all(evaluations == 0)){
+        warning(.msg.invalid.optimization.results.grid)
+        return(filterObj)
+    }
+    optim.pars <- par.grid[which.max(evaluations),]
+    newFilter <- update(filterObj, as.numeric(optim.pars))
+    return(newFilter)
+}
+
+
+
 setMethod("optimize_filter",
             signature(filterObj="MSnIDFilter", msnidObj="MSnID"),
             definition=function(filterObj, msnidObj, fdr.max,
@@ -120,19 +150,24 @@ setMethod("optimize_filter",
     }
     #
     if(method == "Grid"){
-        if(.Platform$OS.type == "unix"){
-            if(is.null(mc.cores))
-                mc.cores <- getOption("mc.cores", 2L)
-            optimFilter <-
-                .optimize_filter.grid.mclapply(filterObj, msnidObj,
-                                               fdr.max, level, 
-                                               n.iter, mc.cores)
-        }else{
-            # yet to implement effective parallelization on Windows
-            optimFilter <-
-                .optimize_filter.grid(filterObj, msnidObj,
-                                        fdr.max, level, n.iter)
-        }
+        # if(.Platform$OS.type == "unix"){
+        #     if(is.null(mc.cores))
+        #         mc.cores <- getOption("mc.cores", 2L)
+        #     optimFilter <-
+        #         .optimize_filter.grid.mclapply(filterObj, msnidObj,
+        #                                        fdr.max, level, 
+        #                                        n.iter, mc.cores)
+        # }else{
[…]

### 2020-07-13 R/AllGenerics.R
Revert "Add argument to enable prioritized inference"

@@ -1,6 +1,6 @@
 
 setGeneric("infer_parsimonious_accessions",
-           function(object, unique_only=FALSE, parsimony_map = NULL) 
+           function(object, unique_only=FALSE) 
                standardGeneric("infer_parsimonious_accessions"))
 
 

### 2020-07-14 R/AllGenerics.R
key PTM handling methods

@@ -1,4 +1,19 @@
 
+setGeneric("add_mod_symbol",
+           function(object, ...)
+               standardGeneric("add_mod_symbol"))
+
+
+setGeneric("report_mods",
+           function(object) 
+               standardGeneric("report_mods"))
+
+
+setGeneric("map_mod_sites",
+           function(object, ...) 
+               standardGeneric("map_mod_sites"))
+
+
 setGeneric("infer_parsimonious_accessions",
            function(object, unique_only=FALSE) 
                standardGeneric("infer_parsimonious_accessions"))

### 2020-07-14 R/MSnID-methods.R

@@ -634,3 +634,74 @@ setMethod("infer_parsimonious_accessions", "MSnID",
 )
 
 
+
+setMethod("report_mods", 
+          signature = signature("MSnID"),
+          definition = function(object)
+          {
+              mods <- psms(object)$modification
+              mod_masses <- lapply(mods, strsplit, "\\s\\(\\d+\\),?\\s?")
+              return(table(unlist(mod_masses)))
+          }
+)
+
+
+setMethod("add_mod_symbol", "MSnID",
+          definition=function(object, mod_mass, symbol)
+          {
+              .add_mod_symbol(object, mod_mass, symbol)
+          }
+)
+
+
+
+
+setMethod("map_mod_sites", "MSnID",
+          definition=function(object, fasta, prot_id_col, peptide_col, mod_char, 
+                              site_delimiter="lower")
+          {
+              ids <- psms(object)
+              
+              # test if decoys are in fasta
+              # if not, then add reversed sequences
+              # for now we'll support only reverse as decoys
+              decoy_acc <- apply_filter(object, "isDecoy")[[prot_id_col]]
+              if(length(decoy_acc) > 0 & !any(decoy_acc %in% names(fasta))){
+                  fasta_rev <- reverse(fasta)
+                  names(fasta_rev) <- paste0("XXX_",names(fasta))
+                  fasta <- c(fasta, fasta_rev)
+              }
+              
+              res <- .map_mod_sites(ids, 
+                                    fasta, 
+                                    prot_id_col, 
+                                    peptide_col, 
+                                    mod_char)
+              
+              # make site ID from prot_id_col (likely accession) and 
+              # SiteCollapsedFirst
+              if(site_delimiter == "lower"){
+                  res <- mutate(res, 
+                                SiteID = paste0(!!sym(prot_id_col), 
+                                                "-", 
+                                                gsub("([[:upper:]])(\\d+),?",
+                                                     "\\1\\2\\L\\1", 
+                                                     SiteCollapsedFirst, 
+                                                     perl = T)))
+              }else{
+                  res <- mutate(res, 
+                                SiteID = paste0(!!sym(prot_id_col), 
+                                                "-", 
+                                                gsub(",",
+                                                     site_delimiter,
+                                                     SiteCollapsedFirst)))
+              }
+              psms(object) <- res
+              return(object)
+          }
+)
+
+
+
+
+

### 2020-07-14 R/add_mod_symbol.R

@@ -1,0 +1,60 @@
+
+rec_insert <- function(x, y, s){
+    if(length(y) == 0)
+        return(x)
+    else{
+        x <- paste0(substr(x, 0, y[1]), s, substr(x, y[1] + 1, nchar(x)))
+        y <- y[-1]
+        rec_insert(x, y, s)
+    }
+}
+
+
+
+.add_mod_symbol <- function(obj, mod_mass, symbol){
+    
+    #
+    # get the positions of PTMs of the specified mass
+    mod_pos <- obj$modification
+    # Bunch of temp vars because I am not ready to use %>% pipe in this package
+    t1 <- strsplit(mod_pos, mod_mass)
+    t2 <- lapply(t1, Filter, f=function(x) grepl("^\\s",x))
+    t3 <- lapply(t2, sub, pattern = "^\\s\\((\\d+).*",replacement = "\\1")
+    t4 <- lapply(t3, as.numeric)
+    t5 <- lapply(t4, `+`, 2) # account for X. at N-term
+    t6 <- lapply(t5, function(..) .. + seq_along(..) - 1) # account for shifts
+    mod_pos <- t6
+    
+    # map to the peptide sequences
+    pep <- obj$peptide_mod
+    if(is.null(pep)){
+        pep <- obj$peptide
+    }else{
+        # how account for already placed mods?
+        # get positions of non-AA symbols
+        prior_mod_pos <- lapply(pep, gregexpr, pattern = "[^ARNDCQEGHILKMFPSTWYV.-]")
+        prior_mod_pos <- lapply(prior_mod_pos, `[[`, 1)
+        prior_mod_pos <- lapply(prior_mod_pos, as.numeric)
+        prior_mod_pos <- lapply(prior_mod_pos, function(..) .. - seq_along(..)) # back to AA indÏ
+        
+        for(i in seq_along(mod_pos)){
+            y <- mod_pos[[i]]
+            d <- prior_mod_pos[[i]]
+            for(j in seq_along(y)){
+                shift <- sum(y[j] >= d)
+                y[j] <- y[j] + shift
+            }
+            mod_pos[[i]] <- y
+        }
+    }
+    
+    # insert the current ones
+    obj$peptide_mod <- sapply(seq_along(pep), function(..) {
+        rec_insert(pep[..], mod_pos[[..]], symbol)})
+    return(obj)
+    
+}
+
+
+
+

### 2020-07-14 R/map_mod_sites.R

@@ -1,0 +1,109 @@
+
+
+# object <- .map_PTM_sites(ids, 
+#                          fasta, 
+#                          prot_id_col, 
+#                          peptide_col, 
+#                          mod_char)
+
+
+# prot_id_col = "accession"
+# peptide_col = "peptide"
+
+.map_mod_sites <- function(ids, fasta, prot_id_col, peptide_col, mod_char){
+    
+    # check if fasta entry names are unique
+    if(any(duplicated(names(fasta)))){
+        message("FASTA entry names are not unique!\n")
+        stop()
+    }
+    
+    # check if there is at least some agreement in IDs
+    if(length(intersect(ids[[prot_id_col]], names(fasta))) == 0){
+        message("There is zero overlap in protein IDs and FASTA entry names!\n")
+        stop()
+    }
+    
+    # check the characters
+    # there should be nothing except the AAs and modification character
+    # ids <- ids %>%
+    #    mutate_(TrimmedPeptide := sub(".\\.(.*)\\..", "\\1", peptide_col))
+    ids <- ids %>%
+        mutate(TrimmedPeptide = sub(".\\.(.*)\\..", "\\1", .[[peptide_col]]))
+    present_chars <- paste0(ids$TrimmedPeptide, collapse = '') %>% 
+        strsplit(split='') %>% 
+        `[[`(1) %>% 
+        unique()
+    other_chars <- setdiff(present_chars, c(AA_STANDARD, mod_char))
+    if(length(other_chars) > 0){
+        message("Detected extra chararacters in the peptide sequences!\n")
+        # erase other chars in TrimmedPeptide
+        other_chars_pttrn <- other_chars %>% 
+            map_chr(~paste0("\\",.x)) %>% 
+            paste0(collapse='') %>% 
+            paste0("[",.,"]")
+        ids <- ids %>%
+            mutate(TrimmedPeptide = str_replace_all(TrimmedPeptide, other_chars_pttrn, ""))
+    }
+    
+    # check for missing peptide sequences
+    if(any(is.na(ids$TrimmedPeptide))){
+        message("Entries with missing peptide sequences were detected. 
+The correspoding identifications will be removed.\n")
+        ids <- ids %>% filter(!is.na(TrimmedPeptide))
+    }
+    
+    # check if there are peptides without mods
[…]

### 2020-07-15 R/MSnID-methods.R
revisited inference method. incorporated "prior" argument.

@@ -578,56 +578,53 @@ setAs("MSnID", "MSnSet",
 
 
 setMethod("infer_parsimonious_accessions", "MSnID",
-          definition=function(object, unique_only=FALSE)
+          definition=function(object, unique_only=FALSE, prior=character(0))
           {
-              # # Old code for inferring accessions.
-              # # It is too slow.
-              # infer_acc <- function(x){
-              #     res <- list()
-              #     while(nrow(x) > 0){
-              #         top_prot <- names(which.max(table(x[['accession']])))
-              #         top_peps <- subset(x, accession == top_prot)
-              #         res <- c(res, list(top_peps))
-              #         x <- subset(x, !(pepSeq %in% top_peps[,"pepSeq"]))
-              #     }
-              #     return(Reduce(rbind,res))
-              # }
-
               infer_acc <- function(x){
                   res <- list()
                   while(nrow(x) > 0){
                       top_prot <- x[, .N, by=accession][which.max(N),,]$accession
                       top_peps <- subset(x, accession == top_prot)
+                      # top_peps <- x[accession == top_prot, c(1,2), with=FALSE] # slower
                       res <- c(res, list(top_peps))
                       x <- subset(x, !(pepSeq %in% top_peps[[1]]))
+                      # x <- x[!top_peps, on=.(pepSeq)] # slower
                   }
                   return(rbindlist(res, use.names=F, fill=FALSE, idcol=NULL))
               }
               
-              x <- unique(psms(object)[,c("pepSeq","accession")])
-              setDT(x)
-              # order, so at least there is certainty which is first
+              x <- unique(object@psms[,.(pepSeq, accession)])
               setorder(x, accession, pepSeq)
+              
               if(unique_only){
                   redundancy <- x[, .(num = length(accession)), by = pepSeq]
                   non_redundant <- redundancy[num == 1]
                   res <- non_redundant[,.(pepSeq)]
                   setkey(res, pepSeq)
-              }else{
-                  # razor
-                  res <- infer_acc(x) # this step may take awhile
+              }else{ # consider razor peptides as well
+                  
+                  # peptides from proteins justified by prior
+                  xp <- x[accession %in% prior][,.(pepSeq)]
+                  xp <- unique(xp)
+                  #
+                  x_prior <- x[xp, on=.(pepSeq)] # semi_join
+                  x_current <- x[!xp, on=.(pepSeq)] # anti_join
+                  
+                  # this removes other proteins mapped to 
+                  # peptides from prior justified proteins
[…]

### 2020-07-17 R/AllGenerics.R
lots of documentation fixes

@@ -1,21 +1,26 @@
 
 setGeneric("add_mod_symbol",
-           function(object, ...)
+           function(object, mod_mass, symbol)
                standardGeneric("add_mod_symbol"))
 
 
 setGeneric("report_mods",
-           function(object) 
+           function(object, ...) 
                standardGeneric("report_mods"))
 
 
 setGeneric("map_mod_sites",
-           function(object, ...) 
+           function(object,
+                    fasta,
+                    accession_col = "accession",
+                    peptide_mod_col = "peptide_mod",
+                    mod_char = "*",
+                    site_delimiter = "lower")
                standardGeneric("map_mod_sites"))
 
 
 setGeneric("infer_parsimonious_accessions",
-           function(object, unique_only=FALSE) 
+           function(object, unique_only=FALSE, prior=character(0)) 
                standardGeneric("infer_parsimonious_accessions"))
 
 

### 2020-07-17 R/MSnID-methods.R

@@ -575,7 +575,7 @@ setAs("MSnID", "MSnSet",
 
 
 
-
+utils::globalVariables(c("accession", "N", "pepSeq", "num"))
 
 setMethod("infer_parsimonious_accessions", "MSnID",
           definition=function(object, unique_only=FALSE, prior=character(0))
@@ -651,18 +651,20 @@ setMethod("add_mod_symbol", "MSnID",
 )
 
 
-
-
 setMethod("map_mod_sites", "MSnID",
-          definition=function(object, fasta, prot_id_col, peptide_col, mod_char, 
-                              site_delimiter="lower")
+          definition=function(object, 
+                              fasta, 
+                              accession_col = "accession", 
+                              peptide_mod_col = "peptide_mod", 
+                              mod_char = "*", 
+                              site_delimiter = "lower")
           {
               ids <- psms(object)
               
               # test if decoys are in fasta
               # if not, then add reversed sequences
               # for now we'll support only reverse as decoys
-              decoy_acc <- apply_filter(object, "isDecoy")[[prot_id_col]]
+              decoy_acc <- apply_filter(object, "isDecoy")[[accession_col]]
               if(length(decoy_acc) > 0 & !any(decoy_acc %in% names(fasta))){
                   fasta_rev <- reverse(fasta)
                   names(fasta_rev) <- paste0("XXX_",names(fasta))
@@ -671,15 +673,15 @@ setMethod("map_mod_sites", "MSnID",
               
               res <- .map_mod_sites(ids, 
                                     fasta, 
-                                    prot_id_col, 
-                                    peptide_col, 
+                                    accession_col, 
+                                    peptide_mod_col, 
                                     mod_char)
               
-              # make site ID from prot_id_col (likely accession) and 
+              # make site ID from accession_col and 
               # SiteCollapsedFirst
               if(site_delimiter == "lower"){
                   res <- mutate(res, 
-                                SiteID = paste0(!!sym(prot_id_col), 
+                                SiteID = paste0(!!sym(accession_col), 
                                                 "-", 
                                                 gsub("([[:upper:]])(\\d+),?",
                                                      "\\1\\2\\L\\1", 
@@ -687,7 +689,7 @@ setMethod("map_mod_sites", "MSnID",
                                                      perl = T)))
               }else{
                   res <- mutate(res, 
-                                SiteID = paste0(!!sym(prot_id_col), 
+                                SiteID = paste0(!!sym(accession_col), 
                                                 "-", 
                                                 gsub(",",
                                                      site_delimiter,

### 2020-07-17 R/map_mod_sites.R

@@ -1,16 +1,15 @@
 
 
-# object <- .map_PTM_sites(ids, 
-#                          fasta, 
-#                          prot_id_col, 
-#                          peptide_col, 
-#                          mod_char)
+utils::globalVariables(c(".", "TrimmedPeptide", "x",
+                         "ProtSeq", "CleanSeq", "PepLoc", "ModShift",
+                         "SiteLoc", "ModAAs", "SiteLoc", "ModAAs",
+                         "Site", "SiteCollapsed", "SiteCollapsedFirst"))
 
-
-# prot_id_col = "accession"
-# peptide_col = "peptide"
-
-.map_mod_sites <- function(ids, fasta, prot_id_col, peptide_col, mod_char){
+.map_mod_sites <- function(ids, 
+                           fasta, 
+                           accession_col, 
+                           peptide_mod_col, 
+                           mod_char){
     
     # check if fasta entry names are unique
     if(any(duplicated(names(fasta)))){
@@ -19,7 +18,7 @@
     }
     
     # check if there is at least some agreement in IDs
-    if(length(intersect(ids[[prot_id_col]], names(fasta))) == 0){
+    if(length(intersect(ids[[accession_col]], names(fasta))) == 0){
         message("There is zero overlap in protein IDs and FASTA entry names!\n")
         stop()
     }
@@ -27,9 +26,9 @@
     # check the characters
     # there should be nothing except the AAs and modification character
     # ids <- ids %>%
-    #    mutate_(TrimmedPeptide := sub(".\\.(.*)\\..", "\\1", peptide_col))
+    #    mutate_(TrimmedPeptide := sub(".\\.(.*)\\..", "\\1", peptide_mod_col))
     ids <- ids %>%
-        mutate(TrimmedPeptide = sub(".\\.(.*)\\..", "\\1", .[[peptide_col]]))
+        mutate(TrimmedPeptide = sub(".\\.(.*)\\..", "\\1", .[[peptide_mod_col]]))
     present_chars <- paste0(ids$TrimmedPeptide, collapse = '') %>% 
         strsplit(split='') %>% 
         `[[`(1) %>% 
@@ -70,9 +69,9 @@ The correspoding identifications will be removed.\n")
     # merger of identifications and FASTA
     res <- fasta %>%
         as.data.frame() %>%
-        rownames_to_column(prot_id_col) %>%
+        rownames_to_column(accession_col) %>%
         rename(ProtSeq = x) %>%
[…]

### 2020-07-17 R/optimize_filter.r

@@ -37,7 +37,7 @@
     eg <- expand.grid(lapply(names(filterObj),
                                 function(arg)
                                     quantile(msnidObj[[arg]],
-                                                probs, na.rm=TRUE)))
+                                             probs, na.rm=TRUE)))
     colnames(eg) <- names(filterObj)
     return(eg)
 }

### 2020-07-17 R/read_mzIDs.R

@@ -1,4 +1,9 @@
 
+utils::globalVariables(c("i", "spectrumID", "name", "mass", "location", 
+                         "modification", "DatabaseAccess", 
+                         "DatabaseDescription", "DBseqLength"))
+
+
 .read_mzIDs.memoized <- function(mzids)
 {
     # Try to load cached data, if exists

### 2020-07-18 R/add_mod_symbol.R
fixed shift when there are no prior modification were mapped

@@ -35,13 +35,14 @@ rec_insert <- function(x, y, s){
         prior_mod_pos <- lapply(pep, gregexpr, pattern = "[^ARNDCQEGHILKMFPSTWYV.-]")
         prior_mod_pos <- lapply(prior_mod_pos, `[[`, 1)
         prior_mod_pos <- lapply(prior_mod_pos, as.numeric)
-        prior_mod_pos <- lapply(prior_mod_pos, function(..) .. - seq_along(..)) # back to AA indÏ
+        prior_mod_pos <- lapply(prior_mod_pos, function(..) ifelse(.. < 0, NA, ..))
+        prior_mod_pos <- lapply(prior_mod_pos, function(..) .. - seq_along(..)) # back to AA ind
         
         for(i in seq_along(mod_pos)){
             y <- mod_pos[[i]]
             d <- prior_mod_pos[[i]]
             for(j in seq_along(y)){
-                shift <- sum(y[j] >= d)
+                shift <- sum(y[j] >= d, na.rm = TRUE)
                 y[j] <- y[j] + shift
             }
             mod_pos[[i]] <- y

### 2020-07-24 R/fetch_conversion_table.R
addition of fetch_conversion_table function

@@ -1,0 +1,62 @@
+
+utils::globalVariables("REFSEQ")
+
+
+fetch_conversion_table <- function(organism_name, from, to, backend="AnnotationHub", snapshot_date=NULL){
+    
+    wrong_column_name <- function(col_name){
+        message(paste0(col_name," is not a valid column name in selected OrgDb."))
+        message("Valid from/to names:")
+        message(paste(AnnotationDbi::columns(db), collapse = ", "))
+        stop()
+    }
+    
+    ah <- AnnotationHub()
+    
+    if(!is.null(snapshot_date)){
+        if(snapshot_date %in% AnnotationHub::possibleDates(ah)){
+            AnnotationHub::snapshotDate(ah) <- snapshot_date
+        }else{
+            message(paste0(snapshot_date, " is not valid snapshot date"))
+            message("valid dates:")
+            message(paste(AnnotationHub::possibleDates(ah), collapse = ", "))
+            stop()
+        }
+    }
+    
+    orgs <- subset(ah, ah$rdataclass == "OrgDb")
+    
+    if(!(organism_name %in% orgs$species)){
+        stop(paste0(organism_name," is not in AnnotationHub database."))
+    }
+    
+    db <- query(orgs, organism_name)
+    db <- db[[1]]
+    
+    if(!(from %in% AnnotationDbi::columns(db))){
+        wrong_column_name(from)
+    }
+    if(!(to %in% AnnotationDbi::columns(db))){
+        wrong_column_name(to)
+    }
+    
+    # main call
+    conversion_table <- AnnotationDbi::select(db,
+                                              keys = AnnotationDbi::keys(db),
+                                              columns = c(from,to),
+                                              keytype = "ENTREZID")
+    conversion_table <- unique(conversion_table[,c(from,to)])
+    conversion_table <- conversion_table[complete.cases(conversion_table),]
+    
+    # rectify the fact that REFSEQ returns both transcripts and proteins
+    # which is a mess
+    if("REFSEQ" %in% c(from, to)){
+        conversion_table <- conversion_table %>%
+            filter(grepl("^.P_.*", REFSEQ))
+    }
+    
+    return(conversion_table)
+}
+
+
+

### 2020-07-24 R/remap_fasta_entry_names.R
adding remaping fasta entry names function

@@ -1,0 +1,57 @@
+
+
+
+
+
+
+remap_fasta_entry_names <- function(path_to_FASTA,
+                                    conversion_table,
+                                    fasta_entry_name_pttrn=c("\\|([^|-]+)(-\\d+)?\\|","^([A-Z]P_\\d+)","^(ENS[A-Z0-9]+)")){
+    
+    is_compressed <- FALSE
+    if(grepl("[.]gz$", path_to_FASTA)){
+        is_compressed <- TRUE
+    }else if(grepl("[.](bz2|xz|zip)$", path_to_FASTA)){
+        stop("The only supported compression is gzip!")
+    }
+    
+    mySequences <- readAAStringSet(path_to_FASTA)
+    fasta_entry_name_pttrn <- paste0(".*", fasta_entry_name_pttrn, ".*")
+    names(mySequences) <- sub(fasta_entry_name_pttrn, "\\1", names(mySequences))
+    prot_lengths <- data.frame(seq_name = names(mySequences),
+                               seq_length = width(mySequences),
+                               stringsAsFactors = FALSE)
+    
+    by_ <- "seq_name"
+    names(by_) <- colnames(conversion_table)[1]
+    conversion_table <- conversion_table %>%
+        inner_join(prot_lengths, by = by_) %>%
+        group_by_at(colnames(conversion_table)[2]) %>%
+        dplyr::slice(which.max(seq_length))
+    
+    # in case there are multiple "to" IDs, the named vector will return
+    # just the first one anyway
+    conversion_table <- conversion_table %>%
+        select(-seq_length) %>%
+        distinct() %>%
+        group_by_at(colnames(conversion_table)[1]) %>%
+        top_n(1,dplyr::desc(!!as.name(colnames(conversion_table)[2])))
+    
+    conversion_vec <- conversion_table %>% pull(2)
+    names(conversion_vec) <- conversion_table %>% pull(1)
+    
+    mySequences <- mySequences[names(conversion_vec)]
+    names(mySequences) <- conversion_vec[names(mySequences)]
+    
+    file_no_ext <- tools::file_path_sans_ext(path_to_FASTA, compression=TRUE)
+    ext <- sub(file_no_ext, "", path_to_FASTA, fixed=TRUE)
+    
+    path_to_FASTA_remapped <- paste0(file_no_ext, '_remapped', ext)
+    
+    writeXStringSet(mySequences, path_to_FASTA_remapped, compress = is_compressed)
+    return(path_to_FASTA_remapped)
+}
+
+
+
+

### 2020-08-05 R/remap_accessions-method.R
small fixed in couple of methods

@@ -33,7 +33,10 @@ setMethod("remap_accessions", "MSnID",
     extraction_pttrn <- paste0(".*", extraction_pttrn, ".*")
     object$accession <- sub(extraction_pttrn, "\\1", object$accession)
     
-    if(prop.table(table(object$accession %in% names(conv_vec)))['FALSE'] > 0.5){
+    # checking that the conversion table covers the accessions
+    cvrg <- object$accession %in% names(conv_vec)
+    cvrg <- factor(cvrg, levels = c(TRUE, FALSE))
+    if(prop.table(table(cvrg))['FALSE'] > 0.5){
         stop("Majority of accessions not in the conversion_table!")
     }
     

### 2020-08-05 R/remap_fasta_entry_names.R

@@ -3,7 +3,7 @@ utils::globalVariables("seq_length")
 
 remap_fasta_entry_names <- function(path_to_FASTA,
                                     conversion_table,
-                                    fasta_entry_name_pttrn=c("\\|([^|-]+)(-\\d+)?\\|","^([A-Z]P_\\d+)","^(ENS[A-Z0-9]+)")){
+                                    extraction_pttrn=c("\\|([^|-]+)(-\\d+)?\\|","^([A-Z]P_\\d+)","^(ENS[A-Z0-9]+)")){
     
     is_compressed <- FALSE
     if(grepl("[.]gz$", path_to_FASTA)){
@@ -13,8 +13,8 @@ remap_fasta_entry_names <- function(path_to_FASTA,
     }
     
     mySequences <- readAAStringSet(path_to_FASTA)
-    fasta_entry_name_pttrn <- paste0(".+?", fasta_entry_name_pttrn, ".+?")
-    names(mySequences) <- sub(fasta_entry_name_pttrn, "\\1", names(mySequences))
+    extraction_pttrn <- paste0(".*", extraction_pttrn, ".*")
+    names(mySequences) <- sub(extraction_pttrn, "\\1", names(mySequences))
     prot_lengths <- data.frame(seq_name = names(mySequences),
                                seq_length = width(mySequences),
                                stringsAsFactors = FALSE)

### 2020-08-06 R/AllGenerics.R
switch to safer ID pattern extraction

@@ -3,8 +3,8 @@ setGeneric("remap_accessions",
            function(object, 
                     conversion_table, 
                     extraction_pttrn=c("\\|([^|-]+)(-\\d+)?\\|",
-                                       "^([A-Z]P_\\d+)",
-                                       "^(ENS[A-Z0-9]+)"), 
+                                       "([A-Z]P_\\d+)",
+                                       "(ENS[A-Z0-9]+)"), 
                     path_to_FASTA=NULL)
                standardGeneric("remap_accessions"))
 

### 2020-08-06 R/remap_fasta_entry_names.R

@@ -3,7 +3,9 @@ utils::globalVariables("seq_length")
 
 remap_fasta_entry_names <- function(path_to_FASTA,
                                     conversion_table,
-                                    extraction_pttrn=c("\\|([^|-]+)(-\\d+)?\\|","^([A-Z]P_\\d+)","^(ENS[A-Z0-9]+)")){
+                                    extraction_pttrn=c("\\|([^|-]+)(-\\d+)?\\|",
+                                                       "([A-Z]P_\\d+)",
+                                                       "(ENS[A-Z0-9]+)")){
     
     is_compressed <- FALSE
     if(grepl("[.]gz$", path_to_FASTA)){

### 2020-08-07 R/remap_accessions-method.R
make sure NA accessions don't make through

@@ -42,6 +42,9 @@ setMethod("remap_accessions", "MSnID",
     
     object$accession <- conv_vec[object$accession]
     
+    # get rid of entries without annotation
+    object <- apply_filter(object, "!is.na(accession)")
+    
     # make sure decoy accessions start with XXX
     object$accession <- ifelse(object$isDecoy,
                               paste0("XXX_", object$accession),
